---
title: "comb_then_split"
author: "Sanmi Adenaiye"
date: "4/14/2021"
output:
  html_document: default
---
Generate datasets for prediction analyses and plots by combining qPCR replicate data with symptom and demographic data.

These sets were developed for the nomask, paired, and nonpaired (combined nomask/yesmask) datasets as separate duplicate scripts. Later they were refined and then recombined into one script.

Revision dates 6/3/21 by Sanmi Adenaiye and 7/6/21-7/10/21 by Filbert Hong
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```
demo_cols, pcr_cols, and sllseq, keysubjects with clade identification
```{r}
#columns needed in demohist for prediction
demo_cols <- c("study_id","real_onset","pos_enrollment","bmi","age_entry","sex_birth","cigarette_ever")

pcr_cols <- c("oral_temp",
              "nose_run","nose_stuf","sneeze","throat_sr","earache","malaise","headache","mj_ache","sw_fever_chill","nausea",
              "loa","vomit","diarrhea","chest_tight","sob","cough",
              "ur","lr","gi","sys",
              "taste_smell_loss","ox_sat","fever_samp","days_past_onset")

allseqs <- read.csv(file="../../source_data/allseqs.csv",stringsAsFactors=F)
keysubjects <- read.csv("../working_files/keysubjectsplus.csv",stringsAsFactors = F)
```
create keysubjects_varseq
```{r}
keysubjects_varseq <- keysubjects %>% left_join(allseqs,by="study_id") %>% select(study_id,clade,Pango.Lineage) %>% unique()
keysubjects_varseq[is.na(keysubjects_varseq)] <- "undetermined"
keysubjects_varseq$study_id <- keysubjects_varseq$study_id %>% as.character()
```
Import nomask data
```{r}
cov_pcr_sera_nomask_complete <- readRDS("../working_files/s1_datasets/s1_cov_pcr_sera_20210805_nomask_complete.RDS")
demohist_nomask_complete <- cov_pcr_sera_nomask_complete[["demohist"]]
pcr_screen_nomask_complete <- cov_pcr_sera_nomask_complete[["pcr_screen"]]
sero1_nomask_complete <- cov_pcr_sera_nomask_complete[["sero1"]]
qpcr_allsamples_nomask_complete <- cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]

pcr_dat_nomask <- qpcr_allsamples_nomask_complete # %>% select(all_of(vars))

pcr_dat_nomask <- pcr_dat_nomask %>% left_join(demohist_nomask_complete %>% select(all_of(demo_cols)), by="study_id")
pcr_dat_nomask <- pcr_dat_nomask %>% left_join(sero1_nomask_complete %>% select(study_id,serum1,serum2,sero_daydiff,seroconvert) %>% unique(),by="study_id")

pcr_dat_nomask %>% group_by(sample_type) %>% summarise(subjects = length(unique(study_id)),samples = n())
```
# nomask_complete
```{r}
pcr_dat_nomask_complete <- pcr_dat_nomask
```
# yesmask_complete
```{r}
cov_pcr_sera_yesmask_complete <- readRDS("../working_files/s1_datasets/s1_cov_pcr_sera_20210805_yesmask_complete.RDS")
demohist_yesmask_complete <- cov_pcr_sera_yesmask_complete[["demohist"]] 
pcr_screen_yesmask_complete <- cov_pcr_sera_yesmask_complete[["pcr_screen"]]
sero1_yesmask_complete <- cov_pcr_sera_yesmask_complete[["sero1"]]
qpcr_allsamples_yesmask_complete <- cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]

pcr_dat_yesmask_complete <- qpcr_allsamples_yesmask_complete # %>% select(vars)
pcr_dat_yesmask_complete <- pcr_dat_yesmask_complete %>% left_join(demohist_yesmask_complete %>% select(all_of(demo_cols)), by="study_id")
pcr_dat_yesmask_complete <- pcr_dat_yesmask_complete%>% left_join(sero1_yesmask_complete %>% select(study_id,serum1,serum2,sero_daydiff,seroconvert) %>% unique(),by="study_id")

head(pcr_dat_yesmask_complete)
```

## Bind both yesmask and nomask to create the full nonpaired set
```{r}
# Simply binding yesmask and nomask and applying unique() gets rid of non-detect replicates that we want to keep.
# Distinguish the replicates (by numbering them with repnum) to prevent this from happening.

pcr_dat_nomask_complete <- pcr_dat_nomask_complete %>% arrange(study_id,sample_date,sample_id,sample_quant)
pcr_dat_yesmask_complete <- pcr_dat_yesmask_complete %>% arrange(study_id,sample_date,sample_id,sample_quant)

nomasksamples <- pcr_dat_nomask_complete$sample_id %>% unique()
pcr_dat_nomask_complete$repnum <- NA
for (i in nomasksamples) {
  j <- pcr_dat_nomask_complete %>% filter(sample_id==i) %>% nrow()
  pcr_dat_nomask_complete$repnum[which(pcr_dat_nomask_complete$sample_id==i)] <- c(1:j)
}

yesmasksamples <- pcr_dat_yesmask_complete$sample_id %>% unique()
pcr_dat_yesmask_complete$repnum <- NA
for (i in yesmasksamples) {
  j <- pcr_dat_yesmask_complete %>% filter(sample_id==i) %>% nrow()
  pcr_dat_yesmask_complete$repnum[which(pcr_dat_yesmask_complete$sample_id==i)] <- c(1:j)
}

pcr_dat_nonpaired <- pcr_dat_nomask_complete %>% bind_rows(pcr_dat_yesmask_complete) %>% unique()
```

```{r}
demohist_nonpaired <- demohist_yesmask_complete %>% bind_rows(demohist_nomask_complete) %>% unique()
demohist_nonpaired %>% group_by(pos_enrollment) %>% summarise(n=length(unique(study_id)))
```
## paired dataset
```{r}
cov_pcr_sera_paired_complete <- readRDS("../working_files/s1_datasets/s1_cov_pcr_sera_20210805_paired_complete.RDS")
demohist_paired_complete <- cov_pcr_sera_paired_complete[["demohist"]]
pcr_screen_paired_complete <- cov_pcr_sera_paired_complete[["pcr_screen"]]
sero1_paired_complete <- cov_pcr_sera_paired_complete[["sero1"]]
qpcr_allsamples_paired_complete <- cov_pcr_sera_paired_complete[["qpcr_allsamples"]]

pcr_dat_paired <- qpcr_allsamples_paired_complete # %>% select(all_of(vars))
pcr_dat_paired <- pcr_dat_paired %>% left_join(demohist_paired_complete %>% select(all_of(demo_cols)), by="study_id")
pcr_dat_paired <- pcr_dat_paired %>% left_join(sero1_paired_complete %>% select(study_id,serum1,serum2,sero_daydiff,seroconvert) %>% unique(),by="study_id")

pcr_dat_paired %>%
  group_by(sample_type) %>%
  summarise(subjects = length(unique(study_id)),
            samples = n())
```
## add symptoms variable to dataframe
nomask rpr and spr
```{r}
pcr_screen1 <- pcr_screen_nomask_complete %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_screen1 <- pcr_screen1 %>% select(all_of(pcr_cols),helpmerge) %>% distinct(helpmerge,.keep_all = T)

pcr_sx_rpr_nomask <- pcr_dat_nomask %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_sx_rpr_nomask <- pcr_sx_rpr_nomask %>% left_join(pcr_screen1, by="helpmerge")

# add lineage assignments
pcr_sx_rpr_nomask <- pcr_sx_rpr_nomask %>% left_join(keysubjects_varseq,by="study_id")

# sample per row
pcr_sx_spr_nomask <- pcr_sx_rpr_nomask %>% distinct(sample_id,.keep_all = T) 

# count number of replicates per sample type
pcr_sx_rpr_nomask %>%
  group_by(sample_type) %>%
  summarize(replicates=n(),
            subjects=length(unique(study_id)))
```

nonpaired rpr and spr
```{r}
# pcr_screen does not have replicate data so unique will work here
pcr_screen <- pcr_screen_nomask_complete %>% bind_rows(pcr_screen_yesmask_complete) %>% unique()

pcr_screen1 <- pcr_screen %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_screen1 <- pcr_screen1 %>% select(all_of(pcr_cols),helpmerge) %>% distinct(helpmerge,.keep_all = T)

pcr_sx_rpr_nonpaired <- pcr_dat_nonpaired %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_sx_rpr_nonpaired <- pcr_sx_rpr_nonpaired %>% left_join(pcr_screen1, by="helpmerge")

# add clade lineage assignments
pcr_sx_rpr_nonpaired <- pcr_sx_rpr_nonpaired %>% left_join(keysubjects_varseq,by="study_id")

# spr from rpr
pcr_sx_spr_nonpaired <- pcr_sx_rpr_nonpaired %>% distinct(sample_id,.keep_all = T) 

pcr_sx_rpr_nonpaired %>% group_by(sample_type) %>% summarize(replicates = n(), subjects =
                                                     length(unique(study_id)))
```

paired rpr and spr
```{r}
pcr_screen1 <- pcr_screen_paired_complete %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_screen1 <- pcr_screen1 %>% select(all_of(pcr_cols),helpmerge) %>% distinct(helpmerge,.keep_all = T)

pcr_sx_rpr_paired <- pcr_dat_paired %>% mutate(helpmerge=paste(study_id,sample_date))
pcr_sx_rpr_paired <- pcr_sx_rpr_paired %>% left_join(pcr_screen1, by="helpmerge")

# add lineage assignments
pcr_sx_rpr_paired <- pcr_sx_rpr_paired %>% left_join(keysubjects_varseq,by="study_id")

# sample per row
pcr_sx_spr_paired <- pcr_sx_rpr_paired %>% distinct(sample_id,.keep_all = T) 

#count number of replicates per sample type
pcr_sx_rpr_paired %>%
  group_by(sample_type) %>%
  summarize(replicates=n(),
            subjects=length(unique(study_id)))
```

write to files
```{r}
write.csv(pcr_sx_rpr_nomask,"../working_files/nomask_pcr_sx_rpr.csv")
write.csv(pcr_sx_spr_nomask,"../working_files/nomask_pcr_sx_spr.csv")

write.csv(pcr_sx_spr_nonpaired,"../working_files/nonpaired_pcr_sx_spr.csv")
write.csv(pcr_sx_rpr_nonpaired,"../working_files/nonpaired_pcr_sx_rpr.csv")
saveRDS(demohist_nonpaired,"../working_files/nonpaired_demohist.RDS")

write.csv(pcr_sx_spr_paired,"../working_files/paired_pcr_sx_spr.csv")
write.csv(pcr_sx_rpr_paired,"../working_files/paired_pcr_sx_rpr.csv")

write.csv(keysubjects_varseq,file="../working_files/keysubjects_varseq.csv",row.names=F)
```
