---
title: "table_s2_subanalysis_for_table3"
author: "Jianyu Lai"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(mvtnorm)
library(lme4)
library(lmec)
library(tidyverse)
source("z_create_matrix.R")
set.seed(42)
```
## data: nonpaired rpr and spr sets filtered for seronegative participants
```{r}
rpr_nonpaired <- read.csv("../working_files/nonpaired_pcr_sx_rpr.csv")
rpr <- rpr_nonpaired %>% filter(pos_enrollment=="Negative")
rpr <- rpr %>% mutate(sgene = factor(S.gene.dropout),
          days_past_onset = as.numeric(days_past_onset),
          dpo = if_else(!is.na(days_past_onset), days_past_onset, -1))
rpr$cens <- rpr$cens %>% as.factor()

spr_nonpaired <- read.csv("../working_files/nonpaired_pcr_sx_spr.csv")
spr <- spr_nonpaired %>% filter(pos_enrollment=="Negative")
```
Using spr to subset the sample_quant values (at sample level) for mt and saliva that will be used as predictors for later analysis.
Create subsets specific to MTS and saliva
```{r}
mt <- spr %>% filter(sample_type == "Midturbinate Swab")
mt <- mt %>% mutate(mt_av_quant = av_quant)
mt <- mt %>% mutate(log_mt = if_else(mt_av_quant==0, 0, log(mt_av_quant)))
mt <- mt %>% select(log_mt, mt_av_quant, study_id, sample_date)
length(unique(mt$study_id))

saliva <- spr %>% filter (sample_type == "Saliva")
saliva <- saliva %>% mutate(sal_av_quant = av_quant)
saliva <- saliva %>% mutate (log_sal = if_else(sal_av_quant==0, 0, log(sal_av_quant)))
saliva <- saliva %>% select (log_sal, sal_av_quant, study_id, sample_date)
length(unique(saliva$study_id))

# Add sample-level mt and saliva values to the rpr so we can use these variables as predictors for later analysis

sal_rpr <- rpr %>% filter(sample_type == "Saliva") %>% left_join(mt, by=c("study_id"="study_id", "sample_date"="sample_date"))
length(unique(sal_rpr$study_id))
length(unique(sal_rpr$sample_id))

mt_rpr <- rpr %>% filter(sample_type == "Midturbinate Swab") %>% left_join(saliva, by=c("study_id"="study_id", "sample_date"="sample_date"))

length(unique(mt_rpr$study_id))
length(unique(mt_rpr$sample_id))
```
median and IQR for mt and saliva sample quantities  
```{r}
# mt
signif(summary(mt_rpr$sample_quant), 2)

# IQR for the log value
mt_rpr <- mt_rpr %>% mutate(log_mt_rpr = if_else(sample_quant==0, 0, log(sample_quant)))
IQR(mt_rpr$log_mt_rpr)
log_mt_IQR <- IQR(mt_rpr$log_mt_rpr) 

# saliva
signif(summary(sal_rpr$sample_quant),2)

# IQR for the log value
sal_rpr <- sal_rpr %>% mutate(log_sal_rpr = if_else(sample_quant==0, 0, log(sample_quant)))
IQR(sal_rpr$log_sal_rpr)
log_sal_IQR <- IQR(sal_rpr$log_sal_rpr) 
```

#LMECs for MT
Z matrix, cluster, cens vector, yL
```{r}
Z_mt_rpr<- z_create_matrix(mt_rpr)
cluster_mt_rpr <- as.numeric(factor(mt_rpr$study_id))
cens_mt_rpr <- mt_rpr$cens
yL_mt_rpr <- mt_rpr$mdfy_sample_quant %>% log()
```

Unadjusted analysis for MT
```{r}
vars_mt_lmec <- c("sgene","log_sal","age_entry","dpo","gi","lr","sys","ur")
est_ci_mt_list <- list()
for (j in vars_mt_lmec) {
  print(paste0("BIVARIATE ANALYSIS FOR mt AND ",j))
  # use lmer() to output the matrices
  lm <- lmer (yL_mt_rpr ~ (1|study_id) + (1|study_id:sample_date) + eval(parse(text=j)), data=mt_rpr)
  x_matrix <- getME(lm, "X")
  fit_ML <- lmec(yL=yL_mt_rpr, cens=cens_mt_rpr, X=x_matrix, Z=Z_mt_rpr, cluster=cluster_mt_rpr, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_mt_list[[j]] <- c(j, est, low, up)
  names(est_ci_mt_list[[j]]) <- c("varname","beta_lmec","ci_low","ci_high")
}
est_ci_mt_df <- bind_rows(est_ci_mt_list)
est_ci_mt_df
```

#LMECs for saliva
Z matrix, cluster, cens vector, yL
```{r}
Z_sal_rpr<- z_create_matrix(sal_rpr)
cluster_sal_rpr <- as.numeric(factor(sal_rpr$study_id))
cens_sal_rpr <- sal_rpr$cens
yL_sal_rpr <- sal_rpr$mdfy_sample_quant %>% log()
```
Unadjusted analysis for sal
```{r}
vars_sal_lmec <- c("sgene","log_mt","age_entry","dpo","gi","lr","sys","ur")
est_ci_sal_list <- list()
for (j in vars_sal_lmec) {
  print(paste0("BIVARIATE ANALYSIS FOR sal AND ",j))
  # use lmer() to output the matrices
  lm <- lmer (yL_sal_rpr ~ (1|study_id) + (1|study_id:sample_date) + eval(parse(text=j)), data=sal_rpr)
  x_matrix <- getME(lm, "X")
  fit_ML <- lmec(yL=yL_sal_rpr, cens=cens_sal_rpr, X=x_matrix, Z=Z_sal_rpr, cluster=cluster_sal_rpr, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_sal_list[[j]] <- c(j, est, low, up)
  names(est_ci_sal_list[[j]]) <- c("varname","beta_lmec","ci_low","ci_high")
}
est_ci_sal_df <- bind_rows(est_ci_sal_list)
est_ci_sal_df
```

create IQR adjustment values
these are multiplied into the power component of an exponential, so essentially 1 = no adjustment
these are needed for table 3, so save to working files
```{r}
IQRadjustlist <- list()

IQRadjustlist[["log_sal"]] <- log_sal_IQR
IQRadjustlist[["log_mt"]] <- log_mt_IQR
IQRadjustlist[["sgene"]] <- 1
IQRadjustlist[["cough_count"]] <- 1
IQRadjustlist[["age_entry"]] <- 10
IQRadjustlist[["dpo"]] <- 1
IQRadjustlist[["gi"]] <- 1
IQRadjustlist[["lr"]] <- 1.2
IQRadjustlist[["sys"]] <- 3
IQRadjustlist[["ur"]] <- 2

saveRDS(IQRadjustlist,file="../working_files/IQRadjustlist.RDS")
```

```{r}
exptransform <- function(x) {
  y <- exp(x) %>% signif(2)
  return(y)
}

est_ci_sal_df$iqradjust <- est_ci_sal_df$varname %>% sapply(function(x) {IQRadjustlist[[x]]})
est_ci_sal_df$beta_1 <- ((est_ci_sal_df$beta_lmec %>% as.numeric())*(est_ci_sal_df$iqradjust)) %>% sapply(exptransform)
est_ci_sal_df$ci_low_1 <- ((est_ci_sal_df$ci_low %>% as.numeric())*(est_ci_sal_df$iqradjust)) %>% sapply(exptransform)
est_ci_sal_df$ci_high_1 <- ((est_ci_sal_df$ci_high %>% as.numeric())*(est_ci_sal_df$iqradjust)) %>% sapply(exptransform)
est_ci_sal_df$predict1 <- est_ci_sal_df[,c("beta_1","ci_low_1","ci_high_1")] %>% apply(1,function(x) {paste0(x["beta_1"]," (",x["ci_low_1"],", ",x["ci_high_1"],")")})

est_ci_mt_df$iqradjust <- est_ci_mt_df$varname %>% sapply(function(x) {IQRadjustlist[[x]]})
est_ci_mt_df$beta_1 <- ((est_ci_mt_df$beta_lmec %>% as.numeric())*(est_ci_mt_df$iqradjust)) %>% sapply(exptransform)
est_ci_mt_df$ci_low_1 <- ((est_ci_mt_df$ci_low %>% as.numeric())*(est_ci_mt_df$iqradjust)) %>% sapply(exptransform)
est_ci_mt_df$ci_high_1 <- ((est_ci_mt_df$ci_high %>% as.numeric())*(est_ci_mt_df$iqradjust)) %>% sapply(exptransform)
est_ci_mt_df$predict1 <- est_ci_mt_df[,c("beta_1","ci_low_1","ci_high_1")] %>% apply(1,function(x) {paste0(x["beta_1"]," (",x["ci_low_1"],", ",x["ci_high_1"],")")})

table_s2_output <- est_ci_mt_df %>% select(varname,predict1) %>% rename(`MT Unadjusted Estimates`=predict1)
table_s2_output <- table_s2_output %>% full_join(est_ci_sal_df %>% select(varname,predict1), by="varname") %>% rename(`Saliva Unadjusted Estimates`=predict1)

tableorder <- c("sgene","age_entry","dpo","log_mt","log_sal","gi","lr","sys","ur")
table_s2_output_1 <- matrix(tableorder,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(table_s2_output_1) <- "varname"
table_s2_output_1 <- table_s2_output_1 %>% left_join(table_s2_output,by="varname")
table_s2_output_1[is.na(table_s2_output_1)] <- "-"
table_s2_output_1$varname[table_s2_output_1$varname=="sgene"] <- "Alpha Variant"
table_s2_output_1$varname[table_s2_output_1$varname=="age_entry"] <- "Age"
table_s2_output_1$varname[table_s2_output_1$varname=="dpo"] <- "Day post-symptom onset"
table_s2_output_1$varname[table_s2_output_1$varname=="log_mt"] <- "Log mid-turbinate swab"
table_s2_output_1$varname[table_s2_output_1$varname=="log_sal"] <- "Log saliva"
table_s2_output_1$varname[table_s2_output_1$varname=="ur"] <- "Upper respiratory symptoms"
table_s2_output_1$varname[table_s2_output_1$varname=="lr"] <- "Lower respiratory symptoms"
table_s2_output_1$varname[table_s2_output_1$varname=="gi"] <- "Gastrointestinal symptoms"
table_s2_output_1$varname[table_s2_output_1$varname=="sys"] <- "Systemic symptoms"
table_s2_output_1

write.csv(table_s2_output_1,file="../table_output/repo_table_s2_alpha_variant_effect.csv",row.names = F)
```