---
title: "table_s7_probability_culture"
author: "Jianyu Lai"
date: "7/19/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(ggpubr)
source("fixRNAscientific.R")
```

## use nonpaired sample per row set filtered for negsero and known culture status
```{r}
spr <- read.csv("../working_files/nonpaired_pcr_sx_spr.csv")
spr <- spr %>% filter(pos_enrollment=="Negative" & culture_status %in% c("positive", "negative"))
spr <- spr %>% mutate(log_av_quant = if_else(av_quant == 0, 0, log(av_quant)),
                      log_10_av_quant = if_else (av_quant == 0, 0, log10(av_quant)),
                      culture = if_else(culture_status=="positive", 1, 0))
table(spr$culture_status, spr$culture)
```

## glm for MTS and saliva
```{r}

# MTS
mt <- spr %>% filter(sample_type=="Midturbinate Swab")
logit_log_10_mt <- glm(culture ~ log_10_av_quant, data = mt, family = "binomial")
summary(logit_log_10_mt)
mt$log_10_prob <- logit_log_10_mt$fitted.values

# saliva
saliva <- spr %>% filter(sample_type=="Saliva")
logit_log_10_saliva <- glm(culture ~ log_10_av_quant, data = saliva, family = "binomial")
summary(logit_log_10_saliva)
saliva$log_10_prob <- logit_log_10_saliva$fitted.values
```

### Subanalysis to find the probability of culture positive for av_quant as log10 of av_quant from 0 to 10
### Estimate the quantum (TCID63 an TCID50) of infection based on mt and saliva cultures
### the latter using log(p/(1-p)) = exp(beta0+beta1*X)
```{r}
# create vector of input average log10 RNA copies
log10_N <- seq(0,10,1)

# mt
beta0_mt <- logit_log_10_mt$coefficients[1]
beta_mt <- logit_log_10_mt$coefficients[2]
prob_mt <- exp(beta0_mt+beta_mt*log10_N)/(1+exp(beta0_mt+beta_mt*log10_N))
prob_mt <- signif(prob_mt,2)

# saliva
beta0_sal <- logit_log_10_saliva$coefficients[1]
beta_sal <- logit_log_10_saliva$coefficients[2]
prob_sal <- exp(beta0_sal+beta_sal*log10_N)/(1+exp(beta0_sal+beta_sal*log10_N))
prob_sal <- signif(prob_sal,2)

prob_pos <- cbind(log10_N, prob_mt, prob_sal)

p_63 <- 1-exp(-1)
p_50 <- 0.5
TCID63_mt <- (log(p_63/(1-p_63))-beta0_mt)/beta_mt
TCID50_mt <- (log(p_50/(1-p_50))-beta0_mt)/beta_mt

TCID63_sal <- (log(p_63/(1-p_63))-beta0_sal)/beta_sal
TCID50_sal <- (log(p_50/(1-p_50))-beta0_sal)/beta_sal

prob_pos
rbind(TCID50_mt, TCID63_mt, TCID50_sal, TCID63_sal)
```

## probability of culture positive for table 2, s7a, s7b values
```{r}
table2df <- read.csv("../working_files/repo_table_2_semiformatted.csv",stringsAsFactors=F)
table7adf <- read.csv("../working_files/repo_table_s7a_semiformatted.csv",stringsAsFactors=F)
table7bdf <- read.csv("../working_files/repo_table_s7b_semiformatted.csv",stringsAsFactors=F)

table2GMvals <- table2df$GM..95..CI. %>% strsplit("\\(") %>% sapply(head,1) %>% as.numeric()
table7aGMvals <- table7adf$GM..95..CI. %>% strsplit("\\(") %>% sapply(head,1) %>% as.numeric()
table7bGMvals <- table7bdf$GM..95..CI. %>% strsplit("\\(") %>% sapply(head,1) %>% as.numeric()

names(table2GMvals) <- table2df$Sample.Type
names(table7aGMvals) <- table7adf$Sample.Type
names(table7bGMvals) <- table7bdf$Sample.Type

sampletypes1 <- names(table2GMvals) %>% subset(!grepl("droplet",tolower(.))) %>% subset(!grepl("mask",tolower(.)))
valuesGM <- sampletypes1 %>% lapply(function(x) {c(table2GMvals[[x]],table7aGMvals[[x]],table7bGMvals[[x]])}) %>% unlist()

table7amaxvals <- table7adf$Maximum.RNA.copies
table7bmaxvals <- table7bdf$Maximum.RNA.copies

names(table7amaxvals) <- table7adf$Sample.Type
names(table7bmaxvals) <- table7bdf$Sample.Type

valuesmaxRNA <- sampletypes1 %>% lapply(function(x) {c(table7amaxvals[[x]],table7bmaxvals[[x]])}) %>% unlist()
values <- c(valuesGM,valuesmaxRNA)
```

## back-calculate 50% probability of a positive culture for MTS and saliva
```{r}
# MTS
p_50_mt <- 10^(-beta0_mt/beta_mt) #781180.7
p_50_mt_format <- signif(p_50_mt,2)

# saliva
p_50_sal <- 10^(-beta0_sal/beta_sal) #5204405
p_50_sal_format <- signif(p_50_sal,2)

p_50_mt_format <- p_50_mt_format %>% fixRNAscientific()
p_50_mt_format
p_50_sal_format <- p_50_sal_format %>% fixRNAscientific()
p_50_sal_format

textoutput <- paste0("The RNA concentration associated with a 50% probability of a positive culture was ",p_50_mt_format," for MTS and ",p_50_sal_format," for saliva (Figure S2).")
textoutput
write(textoutput,file="../table_output/repo_table_s7_50_probability.txt")
```

```{r}
valuesdf <- matrix(values,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(valuesdf) <- "RNAvalues"
valuesdf$Level <- c(rep("Gemoetric mean level",15),rep("Maximum RNA copies",10))
valuesdf$`Sample Type` <- c(sampletypes1 %>% lapply(rep,3) %>% unlist(),sampletypes1 %>% lapply(rep,2) %>% unlist())
valuesdf$Variant <- c(rep(c("Total","Alpha","Other"),length(sampletypes1)),rep(c("Alpha","Other"),length(sampletypes1)))

beta0_mt <- logit_log_10_mt$coefficients[1]
beta_mt <- logit_log_10_mt$coefficients[2]
beta0_sal <- logit_log_10_saliva$coefficients[1]
beta_sal <- logit_log_10_saliva$coefficients[2]

prob_mt <- function(x) {
  y <- exp(beta0_mt+beta_mt*log10(x))/(1+exp(beta0_mt+beta_mt*log10(x)))
  y <- y %>% signif(2) %>% as.character()
  return(y)
}

prob_saliva <- function(x) {
  y <- exp(beta0_sal+beta_sal*log10(x))/(1+exp(beta0_sal+beta_sal*log10(x)))
  y <- y %>% signif(2) %>% as.character()
  return(y)
}

valuesdf$probmt <- valuesdf$RNAvalues %>% sapply(prob_mt)
valuesdf$probsaliva <- valuesdf$RNAvalues %>% sapply(prob_saliva)

valuesdf <- valuesdf %>% select(Level,`Sample Type`,Variant,RNAvalues,probmt,probsaliva)
valuesdf <- valuesdf %>% rename(`RNA copies`=RNAvalues,`Probability based on MT model`=probmt,`Probability based on saliva model`=probsaliva)

valuesdf$`Sample Type`[valuesdf$`Sample Type`=="Phone Swab"] <- "Fomite"
valuesdf$`Sample Type`[valuesdf$`Sample Type`=="G-II Fine Aerosol"] <- "Fine (≤ 5 µm) Aerosol"
valuesdf$`Sample Type`[valuesdf$`Sample Type`=="G-II Coarse Aerosol"] <- "Coarse (> 5 µm) Aerosol"
write.csv(valuesdf,"../table_output/repo_table_s7_probability_culture.csv",row.names=F)
```
