---
title: "table_3_part1_pred_nonpaired"
author: "Jianyu Lai"
date: "6/20/2021"
output:
  html_document: default
---
analysis of shedding/variants
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(mvtnorm)
library(lme4)
library(lmec)
library(tidyverse)
source("z_create_matrix.R")
set.seed(42)
```

Data: nomask rpr and spr files, filtered for seronegative participants
```{r}
rpr_nomask <- read.csv("../working_files/nomask_pcr_sx_rpr.csv")
rpr <- rpr_nomask %>% filter(pos_enrollment=="Negative")

rpr <- rpr %>% mutate(sgene = factor(S.gene.dropout),
                      days_past_onset = as.numeric(days_past_onset),
                      dpo = if_else(!is.na(days_past_onset), days_past_onset, -1))

spr_nomask <- read.csv("../working_files/nomask_pcr_sx_spr.csv")
spr <- spr_nomask %>% filter(pos_enrollment=="Negative")
```
spr to subset the sample_quant values (at sample level) for mt and saliva that will be used as predictors for later analysis
```{r}
mt <- spr %>% filter(sample_type == "Midturbinate Swab")
mt <- mt %>% mutate(mt_av_quant = av_quant,
                    log_mt = if_else(mt_av_quant==0, 0, log(mt_av_quant))) 
mt <- mt %>% select(log_mt, mt_av_quant, study_id, sample_date)

saliva <- spr %>% filter(sample_type == "Saliva") 
saliva <- saliva %>% mutate(sal_av_quant = av_quant,
                             log_sal = if_else(sal_av_quant==0, 0, log(sal_av_quant)))
saliva <- saliva %>% select(log_sal, sal_av_quant, study_id, sample_date)
```
Use left_join on rpr to add mt and saliva values (at sample level) to the three datasets so we can use these variables as predictors for later analysis
```{r}
fine <- rpr %>% filter(sample_type == "G-II Fine Aerosol") %>%
  left_join(mt, by=c("study_id"="study_id", "sample_date"="sample_date")) %>%
  left_join(saliva, by=c("study_id"="study_id", "sample_date"="sample_date")) 
length(unique(fine$study_id))
length(unique(fine$sample_id))

coarse <- rpr %>% filter(sample_type == "G-II Coarse Aerosol") %>%
  left_join(mt, by=c("study_id"="study_id", "sample_date"="sample_date")) %>% 
  left_join(saliva, by=c("study_id"="study_id", "sample_date"="sample_date")) 

length(unique(coarse$study_id))
length(unique(coarse$sample_id))
```

fine aerosol LMEC
```{r}
Z_fine <- z_create_matrix(fine)
cluster_fine <- as.numeric(factor(fine$study_id))
cens_fine <- fine$cens
yL_fine <- fine$mdfy_sample_quant %>% log()
```
BIVARIATE ANALYSIS FOR FINE AEROSOL
```{r}
vars_fine <- c('sgene', 'log_mt', 'log_sal', 'cough_count', 'gi', 'lr', 'sys', 'ur', 'dpo', 'age_entry')
est_ci_fine_list <- list()

for (i in vars_fine) {
  print(paste0("BIVARIATE ANALYSIS FOR FINE AEROSOL AND ",i))
  lm <- lmer (yL_fine ~ (1|study_id) + (1|study_id:sample_date) + eval(parse(text=i)), data=fine)
  x_matrix <- getME(lm, "X")
  fit_ML = lmec(yL=yL_fine, cens=cens_fine, X=x_matrix, Z=Z_fine, cluster=cluster_fine, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_fine_list[[i]] <- c(i, est, low, up)
  names(est_ci_fine_list[[i]]) <- c("varname","beta_lmec","ci_low","ci_high")
}
est_ci_fine_list_df <- bind_rows(est_ci_fine_list)

# confounders for sgene
vars_fine_sgene_confounders <- vars_fine %>% subset(!(.=="sgene"))
est_ci_fine_sgene_list <- list()
for (j in vars_fine_sgene_confounders) {
  print(paste0("Fine & sgene: ",j))
  lm <- lmer (yL_fine ~ (1|study_id) + (1|study_id:sample_date) + sgene + eval(parse(text=j)), data=fine)
  x_matrix <- getME(lm, "X")
  fit_ML = lmec(yL=yL_fine, cens=cens_fine, X=x_matrix, Z=Z_fine, cluster=cluster_fine, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_fine_sgene_list[[j]] <- c(paste0("sgene_adjusted_for_",j), est, low, up)
  names(est_ci_fine_sgene_list[[j]]) <- c("varname","beta_lmec","ci_low","ci_high")
}

est_ci_fine_sgene_list_df <- bind_rows(est_ci_fine_sgene_list)

est_ci_fine_df <- bind_rows(est_ci_fine_list_df,est_ci_fine_sgene_list_df)
```
Multivariable analysis for Fine aerosol
```{r}
est_ci_fine_multi <- c()
lm <- lmer (yL_fine ~ (1|study_id) + (1|study_id:sample_date) + sgene + cough_count + log_mt + log_sal + gi + sys + ur + age_entry, data=fine)
x_matrix <- getME(lm, "X")

#Run the LEMC()
fit_ML = lmec(yL=yL_fine, cens=cens_fine, X=x_matrix, Z=Z_fine, cluster=cluster_fine, maxstep = 200, varstruct = "unstructured", method = "ML")

vars <- c('intercept', 'sgene', 'cough_count', 'log_mt', 'log_sal', 'gi', 'sys', "ur", "age_entry")
for (i in c(2:9)){
                est <- fit_ML$beta[i]
                se <- sqrt(diag(fit_ML$varFix))[i]
                up <- est+1.96*se
                low <- est-1.96*se
                est_ci <- c(vars[i], est, low, up)
                est_ci_fine_multi <- rbind(est_ci_fine_multi, est_ci)
}
colnames(est_ci_fine_multi) <- c("varname","beta_lmec","ci_low","ci_high")
est_ci_fine_multi
```

coarse aerosol LMEC
```{r}
Z_coarse<- z_create_matrix(coarse)
cluster_coarse <- as.numeric(factor(coarse$study_id))
cens_coarse <- coarse$cens
yL_coarse <- coarse$mdfy_sample_quant %>% log()
```
BIVARIATE ANALYSIS FOR COARSE AEROSOL
```{r}
vars_coarse <- c('sgene', 'log_mt', 'log_sal', 'cough_count', 'gi', 'lr', 'sys', 'ur', 'dpo', 'age_entry')
est_ci_coarse_list <- list()

for (i in vars_coarse) {
  print(paste0("BIVARIATE ANALYSIS FOR COARSE AEROSOL AND ",i))
  lm <- lmer (yL_coarse ~ (1|study_id) + (1|study_id:sample_date) + eval(parse(text=i)), data=coarse)
  x_matrix <- getME(lm, "X")
  fit_ML = lmec(yL=yL_coarse, cens=cens_coarse, X=x_matrix, Z=Z_coarse, cluster=cluster_coarse, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_coarse_list[[i]] <- c(i, est, low, up)
  names(est_ci_coarse_list[[i]]) <- c("varname","beta_lmec","ci_low","ci_high")
}
est_ci_coarse_list_df <- bind_rows(est_ci_coarse_list)

# confounders for sgene
vars_coarse_sgene_confounders <- vars_coarse %>% subset(!(.=="sgene"))
est_ci_coarse_sgene_list <- list()
for (j in vars_coarse_sgene_confounders) {
  print(paste0("Coarse & sgene: ",j))
  lm <- lmer (yL_coarse ~ (1|study_id) + (1|study_id:sample_date) + sgene + eval(parse(text=j)), data=coarse)
  x_matrix <- getME(lm, "X")
  fit_ML = lmec(yL=yL_coarse, cens=cens_coarse, X=x_matrix, Z=Z_coarse, cluster=cluster_coarse, maxstep = 200, varstruct = "unstructured", method = "ML")
  est <- fit_ML$beta[2]
  se <- sqrt(diag(fit_ML$varFix))[2]
  up <- est+1.96*se
  low <- est-1.96*se
  est_ci_coarse_sgene_list[[j]] <- c(paste0("sgene_adjusted_for_",j), est, low, up)
  names(est_ci_coarse_sgene_list[[j]]) <- c("varname","beta_lmec","ci_low","ci_high")
}

est_ci_coarse_sgene_list_df <- bind_rows(est_ci_coarse_sgene_list)

est_ci_coarse_df <- bind_rows(est_ci_coarse_list_df,est_ci_coarse_sgene_list_df)
```
Multivariable analysis for coarse aerosol
```{r}
est_ci_coarse_multi <- c()
lm <- lmer (yL_coarse ~ (1|study_id) + (1|study_id:sample_date) + sgene + cough_count + log_mt + log_sal + gi + lr + sys, data=coarse)
x_matrix <- getME(lm, "X")

#Run the LEMC()
fit_ML = lmec(yL=yL_coarse, cens=cens_coarse, X=x_matrix, Z=Z_coarse, cluster=cluster_coarse, maxstep = 200, varstruct = "unstructured", method = "ML")

vars <- c('intercept', 'sgene', 'cough_count', 'log_mt', 'log_sal', 'gi', 'lr','sys')
for (i in c(2:8)){
                est <- fit_ML$beta[i]
                se <- sqrt(diag(fit_ML$varFix))[i]
                up <- est+1.96*se
                low <- est-1.96*se
                est_ci <- c(vars[i], est, low, up)
                est_ci_coarse_multi <- rbind(est_ci_coarse_multi, est_ci)
}
colnames(est_ci_coarse_multi) <- c("varname","beta_lmec","ci_low","ci_high")
print(est_ci_coarse_multi)

```

IQRadjustlist was created in the Table S2 script
```{r}
IQRadjustlist <- readRDS(file="../working_files/IQRadjustlist.RDS")

exptransform <- function(x) {
  y <- exp(x) %>% signif(2)
  return(y)
}

# fine/coarse unadjusted/unadjusted mean & CIs

meanCIlist <- list(est_ci_fine_df,est_ci_fine_multi,est_ci_coarse_df,est_ci_coarse_multi)
names(meanCIlist) <- c("fine_unadj","fine_adj","coarse_unadj","coarse_adj")

meanCIlist <- meanCIlist %>% lapply(as.data.frame,stringsAsFactors=F)
for (i in names(meanCIlist)) {
  meanCIlist[[i]]$iqradjust <- meanCIlist[[i]]$varname %>% lapply(function(x) {IQRadjustlist[[x]]}) %>% lapply(function(x) {if (length(x)==0) {1} else {x}}) %>% unlist()
  meanCIlist[[i]]$beta_1 <- ((meanCIlist[[i]]$beta_lmec %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$ci_low_1 <- ((meanCIlist[[i]]$ci_low %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$ci_high_1 <- ((meanCIlist[[i]]$ci_high %>% as.numeric())*meanCIlist[[i]]$iqradjust) %>% sapply(exptransform)
  meanCIlist[[i]]$predict1 <- meanCIlist[[i]][,c("beta_1","ci_low_1","ci_high_1")] %>% apply(1,function(x) {paste0(x["beta_1"]," (",x["ci_low_1"],", ",x["ci_high_1"],")")})
}

# IQR/range adjustments come from repo_table_s2

# for unadjusted models, don't need to adjusted entries
meanCIlist[["fine_unadj"]] <- meanCIlist[["fine_unadj"]] %>% filter(!grepl("adjusted",varname))
meanCIlist[["coarse_unadj"]] <- meanCIlist[["coarse_unadj"]] %>% filter(!grepl("adjusted",varname))

table3output <- meanCIlist[["coarse_unadj"]] %>% select(varname,predict1) %>% rename(`Coarse Unadjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["coarse_adj"]] %>% select(varname,predict1),by="varname") %>% rename(`Coarse Adjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["fine_unadj"]] %>% select(varname,predict1),by="varname") %>% rename(`Fine Unadjusted`=predict1)
table3output <- table3output %>% left_join(meanCIlist[["fine_adj"]] %>% select(varname,predict1),by="varname") %>% rename(`Fine Adjusted`=predict1)
table3output[is.na(table3output)] <- "-"

tableorder <- c("sgene","age_entry","dpo","log_mt","log_sal","cough_count","ur","lr","gi","sys")

table3output1 <- matrix(tableorder,ncol=1) %>% as.data.frame(stringsAsFactors=F)
colnames(table3output1) <- "varname"
table3output1 <- table3output1 %>% left_join(table3output,by="varname")

table3output1

write.csv(table3output1,file="../table_output/repo_table_3_part1_temp.csv",row.names=F)
```
