---
title: "cov_pcr_sera_subsets"
author: "Filbert Hong"
date: "4/21/2021"
output: html_document
---
Subset cov_pcr_sera for participants with complete masked and/or nonmasked sample sets

1. All available data
2. Complete quantitative PCR for samples from the visit: mid-turbinate swabs, saliva, phone swab, coarse, and fine aerosol
Some salivas had no qPCR run on them because their screening result was nondetect. Impute missing samples
3. Complete quantitatuve PCR and complete culture for all samples from the visit.

```{r setup, include=FALSE}
library(dplyr)
```

For dataset 1, filter qPCR_allsamples to exclude masked samples. Filter pcr_screen to exclude visits that don't have any unmasked sessions.
```{r read files}
allsamples <- read.csv("../source_data/StopCOVIDsamples.csv",colClasses = "character") %>% select(study_id, sample_type, sample_id, date_collected_sg) %>%
                rename(sample_date=date_collected_sg)
cov_pcr_sera <- readRDS("cov_pcr_sera_20210805.RDS")
# fix allsamples sample type names
allsamples$sample_type <- allsamples$sample_type %>% gsub(" Derivative","",.)
allsamples$sample_type[allsamples$sample_type=="G2 Fine Aerosol Wash"] <- "G-II Fine Aerosol"
allsamples$sample_type[allsamples$sample_type=="G2 5-Micron Impactor"] <- "G-II Coarse Aerosol"
allsamples$sample_type[allsamples$sample_type=="IcePac Sample"] <- "IcePac"
allsamples$helper <- allsamples$study_id %>% paste(allsamples$sample_date)
```
subset into mask and nomask
use person-sampledate helpers
```{r mask and nomask}
cov_pcr_sera[["qpcr_allsamples"]] <- cov_pcr_sera[["qpcr_allsamples"]] %>% mutate(helper=paste(study_id,sample_date))
cov_pcr_sera[["pcr_screen"]] <- cov_pcr_sera[["pcr_screen"]] %>% mutate(helper=paste(study_id,sample_date))
visit_helpers_all <- cov_pcr_sera[["qpcr_allsamples"]] %>% pull(helper) %>% unique()
print(paste0("cov_pcr_sera qpcr helpers: ",cov_pcr_sera[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera rows:")
cov_pcr_sera %>% sapply(nrow)

# identify mask and nomask helpers based on qpcr_allsamples
# remove screening PCR visits from cov_pcr_sera_nomask$pcr_screen that aren't on the nomask visit list
# a lot of them go away because many of the pcr_screen results are driveby salivas from arm 1
# remove demohist and sero subject IDs that aren't in qpcr_allsamples$study_id for each respective set

visit_helpers_nomask <- cov_pcr_sera[["qpcr_allsamples"]] %>% filter(mask_yn==0) %>% pull(helper) %>% unique()
visit_helpers_yesmask <- cov_pcr_sera[["qpcr_allsamples"]] %>% filter(mask_yn==1) %>% pull(helper) %>% unique()

cov_pcr_sera_nomask <- cov_pcr_sera
cov_pcr_sera_nomask[["qpcr_allsamples"]] <- cov_pcr_sera_nomask[["qpcr_allsamples"]] %>% filter(helper %in% visit_helpers_nomask) %>% filter(mask_yn!=1)
cov_pcr_sera_nomask[["pcr_screen"]] <- cov_pcr_sera_nomask[["pcr_screen"]] %>% filter(helper %in% visit_helpers_nomask)
cov_pcr_sera_nomask[["demohist"]] <- cov_pcr_sera_nomask[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_nomask[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_nomask[["sero1"]] <- cov_pcr_sera_nomask[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_nomask[["qpcr_allsamples"]]$study_id)
print(paste0("cov_pcr_sera_nomask qpcr helpers: ",cov_pcr_sera_nomask[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_nomask rows:")
cov_pcr_sera_nomask %>% sapply(nrow)

cov_pcr_sera_yesmask <- cov_pcr_sera
cov_pcr_sera_yesmask[["qpcr_allsamples"]] <- cov_pcr_sera_yesmask[["qpcr_allsamples"]] %>% filter(helper %in% visit_helpers_yesmask) %>% filter(mask_yn!=0)
cov_pcr_sera_yesmask[["pcr_screen"]] <- cov_pcr_sera_yesmask[["pcr_screen"]] %>% filter(helper %in% visit_helpers_yesmask)
cov_pcr_sera_yesmask[["demohist"]] <- cov_pcr_sera_yesmask[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_yesmask[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_yesmask[["sero1"]] <- cov_pcr_sera_yesmask[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_yesmask[["qpcr_allsamples"]]$study_id)
print(paste0("cov_pcr_sera_yesmask qpcr helpers: ",cov_pcr_sera_yesmask[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_yesmask rows:")
cov_pcr_sera_yesmask %>% sapply(nrow)

excluded_helpers <- visit_helpers_all %>% subset(!(. %in% c(visit_helpers_nomask,visit_helpers_yesmask))) 
# There are person-days for which the samples were quantitated, but they didn't go into the G-II that day. (like subject 56 on 6/6, a contact initial). Those are excluded.
```
For dataset 2, make sure visits with Saliva, MT, G-II Fine and Coarse Aerosol, Cone Swab, and Phone are present. 

For nomask.

- some visits were done before Cone Swab procedure was implemented. Make sure these visits aren't excluded
- some salivas had no qPCR run on them because their screening result was nondetect. Impute these as rxn_quant=0.
- 274 2020-10-01 saliva was accidentally discarded in between screening and quantitation. Impute this one as rxn_quant=estimated from pcr_screen cts via an average of all std3fits.
```{r complete qPCR set for nomask}
cov_pcr_sera_nomask_complete <- cov_pcr_sera_nomask

visit_helpers_nomask_sampletypes <- visit_helpers_nomask %>% lapply(function(x) {cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] %>% filter(helper==x) %>% select(sample_type) %>% unlist() %>% unique()}) # what samples are in which visit
names(visit_helpers_nomask_sampletypes) <- visit_helpers_nomask

fullset1nomask <- c("Saliva","G-II Coarse Aerosol","G-II Fine Aerosol","Phone Swab","Midturbinate Swab","G-II Cone Swab") # the ideal full set
fullset1anomask <- c("Saliva","G-II Coarse Aerosol","G-II Fine Aerosol","Phone Swab","Midturbinate Swab") # for comparison if the G-II cone was missing
fullset2nomask <- c("G-II Coarse Aerosol","G-II Fine Aerosol","Phone Swab","Midturbinate Swab") # for comparison if the saliva was missing

visit_helpers_nomask_fullset1 <- visit_helpers_nomask_sampletypes %>% subset(lapply(.,function(x) {all(fullset1nomask %in% x)}) %>% unlist()) %>% names() # these are the sets that have full sets
fullset1ahelpers_nomask <- visit_helpers_nomask_sampletypes %>% subset(lapply(.,function(x) {all(fullset1anomask %in% x)}) %>% unlist()) %>% names() # these are the sets that would otherwise be complete except for the Cone Swab
missingconeswab_nomask <- fullset1ahelpers_nomask %>% subset(!(. %in% visit_helpers_nomask_fullset1)) # these sets are the ones who are missing the coneswab
missingconeswabdoesntexist_nomask <- missingconeswab_nomask %>% sapply(function(x) {allsamples %>% filter(helper==x & grepl("cone",tolower(sample_type))) %>% nrow() %>% `==`(0)}) %>% names()
visit_helpers_nomask_fullset1 <- c(visit_helpers_nomask_fullset1,missingconeswabdoesntexist_nomask) # add these helpers back

visit_helpers_nomask_incompl <- visit_helpers_nomask %>% subset(!(. %in% visit_helpers_nomask_fullset1))

cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen <- NA

possiblyincomplete_due_to_saliva_nomask <- visit_helpers_nomask_incompl %>% subset(visit_helpers_nomask_sampletypes[.] %>% sapply(function(x) {all(fullset2nomask %in% x)})) # which ones are missing saliva
incomplete_dueto_saliva_nomask <- possiblyincomplete_due_to_saliva_nomask %>% subset(possiblyincomplete_due_to_saliva_nomask %>% sapply(function(x) {cov_pcr_sera_nomask_complete[["pcr_screen"]] %>% filter(helper==x & grepl("saliva",tolower(sample_type))) %>% select(ct_sgene_screen,ct_ngene_screen,ct_orf1ab_screen) %>% unlist() %>% `==`(c(40,40,40)) %>% all()})) # which ones have saliva nondetects at screening
incomplete_nomask <- visit_helpers_nomask_incompl %>% subset(!(. %in% incomplete_dueto_saliva_nomask))
  
cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen[cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper %in% incomplete_dueto_saliva_nomask] <- T

# manually impute missing salivas
impute_nomask_helpers <- c(incomplete_dueto_saliva_nomask)
impute_nomask_df <- matrix(ncol=ncol(cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]),nrow=length(impute_nomask_helpers)) %>% as.data.frame(stringsAsFactors=F)
colnames(impute_nomask_df) <- colnames(cov_pcr_sera_nomask_complete[["qpcr_allsamples"]])
impute_nomask_df$helper <- impute_nomask_helpers
impute_nomask_df$ct_ngene_qpcr <- -1
impute_nomask_df$ct_sgene_qpcr <- -1
impute_nomask_df$ct_orfgene_qpcr <- -1
impute_nomask_df$nd_lod_loq <- 0
impute_nomask_df$cens <-  1 # used to be logical, now numeric; cens 1 means sample_quant is 0
impute_nomask_df$rxn_quant <- 0
impute_nomask_df$sample_quant <- 0
impute_nomask_df$mdfy_sample_quant <- 75
impute_nomask_df$any_rep_lod1 <- F
impute_nomask_df$all_rep_lod2 <- F
impute_nomask_df$all_rep_loq <- F
impute_nomask_df$av_quant <- 0
impute_nomask_df$loq <- F
impute_nomask_df$S.gene.dropout <- T # yes, this is true for the particular study IDs in this list

# manually impute 274 2020-10-01
helper27420201001 <- c("274 2020-10-01")
names(helper27420201001) <- "helper"
impute_nomask_df <- impute_nomask_df %>% bind_rows(helper27420201001)
impute_nomask_df$ct_ngene_qpcr[impute_nomask_df$helper=="274 2020-10-01"] <- cov_pcr_sera_nomask[["pcr_screen"]] %>% filter(helper=="274 2020-10-01") %>% pull(ct_ngene_screen)
impute_nomask_df$ct_sgene_qpcr[impute_nomask_df$helper=="274 2020-10-01"] <- cov_pcr_sera_nomask[["pcr_screen"]] %>% filter(helper=="274 2020-10-01") %>% pull(ct_sgene_screen)
impute_nomask_df$ct_orfgene_qpcr[impute_nomask_df$helper=="274 2020-10-01"] <- cov_pcr_sera_nomask[["pcr_screen"]] %>% filter(helper=="274 2020-10-01") %>% pull(ct_orf1ab_screen)
# estimate the sample_quant from ct_ngene_qpcr, average of fits
std3_allfits <- read.csv(file="../source_data/std3_allfits.csv",stringsAsFactors=F)
std3_allfits$est_rxn_quant <- ((impute_nomask_df$ct_ngene_qpcr[impute_nomask_df$helper=="274 2020-10-01"]-std3_allfits$b)/std3_allfits$m) %>% `^`(10,.)
impute_nomask_df$rxn_quant[impute_nomask_df$helper=="274 2020-10-01"] <- std3_allfits$est_rxn_quant %>% mean()
impute_nomask_df$sample_quant[impute_nomask_df$helper=="274 2020-10-01"] <- impute_nomask_df$rxn_quant[impute_nomask_df$helper=="274 2020-10-01"] * (125/6)
impute_nomask_df$nd_lod_loq[impute_nomask_df$helper=="274 2020-10-01"] <- 2
impute_nomask_df$cens[impute_nomask_df$helper=="274 2020-10-01"] <-  0 # used to be logical, now numerical; cens 0 means sample_quant >0

impute_nomask_df$any_rep_lod1[impute_nomask_df$helper=="274 2020-10-01"]  <- T
impute_nomask_df$all_rep_lod2[impute_nomask_df$helper=="274 2020-10-01"]  <- T
impute_nomask_df$all_rep_loq[impute_nomask_df$helper=="274 2020-10-01"]  <- T
impute_nomask_df$av_quant[impute_nomask_df$helper=="274 2020-10-01"]  <- impute_nomask_df$sample_quant[impute_nomask_df$helper=="274 2020-10-01"]
impute_nomask_df$mdfy_sample_quant[impute_nomask_df$helper=="274 2020-10-01"]  <- impute_nomask_df$sample_quant[impute_nomask_df$helper=="274 2020-10-01"]
impute_nomask_df$loq[impute_nomask_df$helper=="274 2020-10-01"]  <- T
impute_nomask_df$S.gene.dropout[impute_nomask_df$helper=="274 2020-10-01"]  <- F # this is not an S-gene dropout

impute_nomask_df$study_id <- impute_nomask_df$helper %>% strsplit(" ") %>% lapply(head,1) %>% unlist()
impute_nomask_df$sample_date <- impute_nomask_df$helper %>% strsplit(" ") %>% lapply(tail,1) %>% unlist() %>% as.Date(origin="1970-01-01")
impute_nomask_df$sample_id <- c(1:nrow(impute_nomask_df)) %>% paste0("imputed_saliva_nomask_",.)
impute_nomask_df$exp_name <- "imputed saliva row"
impute_nomask_df$sample_type <- "Saliva"
impute_nomask_df$culture_status <- "imputed sample"
impute_nomask_df$exp_date <- impute_nomask_df$sample_date
impute_nomask_df$gii_sessnum <- "not a G-II sample"
impute_nomask_df$mask_yn <- "not a G-II sample"
impute_nomask_df$mask_type <- "not a G-II sample"
impute_nomask_df$cough_count <-  -999
impute_nomask_df$reserv_time <- -999
impute_nomask_df$gii <- 0 %>% as.factor()
impute_nomask_df$gii_reps <- impute_nomask_df$helper %>% sapply(function(x) {cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] %>% filter(helper==x) %>% pull(gii_reps) %>% unique()})
impute_nomask_df$nd_lod_loq <- impute_nomask_df$nd_lod_loq %>% as.factor()
impute_nomask_df$saliva_missing_after_screen <-  T

cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] <- cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] %>% filter((helper %in% visit_helpers_nomask_fullset1) | (saliva_missing_after_screen) | helper=="274 2020-10-01")
cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] <- cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] %>% bind_rows(impute_nomask_df)

helpers1 <- cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper %>% unique() %>% sort()
if (all(visit_helpers_nomask %in% cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper)) {
  print("all nomask sets are considered quantitated")
} else {
  helpers2 <- visit_helpers_nomask %>% subset(!(. %in% cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper))
  cat(paste0("The following sets are incomplete:\n",helpers2 %>% paste(collapse="\n"),"\n"))
  for (i in 1:length(helpers2)) {
    j <- visit_helpers_nomask_sampletypes[helpers2[i]] %>% unlist()
    notj <- fullset1nomask %>% subset(!(. %in% j))
    cat(paste0(helpers2[i]," has ",j %>% paste(collapse=";"),"; but it's missing ",notj %>% paste(collapse = ",")))
  }
}

# now that cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] has subsetted samples, filter demo1 and sero1 to exclude subjects that are missing, and filter pcr_screen to exclude visits that are not in qpcr_allsamples
cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen[is.na(cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen)] <- F

cov_pcr_sera_nomask_complete[["demohist"]] <- cov_pcr_sera_nomask_complete[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_nomask_complete[["sero1"]] <- cov_pcr_sera_nomask_complete[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_nomask_complete[["pcr_screen"]] <- cov_pcr_sera_nomask_complete[["pcr_screen"]] %>% filter(helper %in% cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper)
print(paste0("cov_pcr_sera_nomask_complete qpcr helpers: ",cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_nomask_complete rows:")
cov_pcr_sera_nomask_complete %>% sapply(nrow)
```
complete sets for yesmask
```{r complete qPCR set for yesmask}
cov_pcr_sera_yesmask_complete <- cov_pcr_sera_yesmask

visit_helpers_yesmask_sampletypes <- visit_helpers_yesmask %>% lapply(function(x) {cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] %>% filter(helper==x) %>% select(sample_type) %>% unlist() %>% unique()}) # what samples are in which visit
names(visit_helpers_yesmask_sampletypes) <- visit_helpers_yesmask

fullset1yesmask <- c("Saliva","G-II Coarse Aerosol mask","G-II Fine Aerosol mask","Phone Swab","Midturbinate Swab","G-II Cone Swab mask") # the ideal full set
fullset1ayesmask <- c("Saliva","G-II Coarse Aerosol mask","G-II Fine Aerosol mask","Phone Swab","Midturbinate Swab") # for comparison if the G-II cone was missing
fullset2yesmask <- c("G-II Coarse Aerosol mask","G-II Fine Aerosol mask","Phone Swab","Midturbinate Swab") # for comparison if the saliva was missing

visit_helpers_yesmask_fullset1 <- visit_helpers_yesmask_sampletypes %>% subset(lapply(.,function(x) {all(fullset1yesmask %in% x)}) %>% unlist()) %>% names() # these are the sets that have full sets
fullset1ahelpers_yesmask <- visit_helpers_yesmask_sampletypes %>% subset(lapply(.,function(x) {all(fullset1ayesmask %in% x)}) %>% unlist()) %>% names() # these are the sets that would otherwise be complete except for the Cone Swab
missingconeswab_yesmask <- fullset1ahelpers_yesmask %>% subset(!(. %in% visit_helpers_yesmask_fullset1)) # these sets are the ones who are missing the coneswab
missingconeswabdoesntexist_yesmask <- missingconeswab_yesmask %>% sapply(function(x) {allsamples %>% filter(helper==x & grepl("cone",tolower(sample_type))) %>% nrow() %>% `==`(0)}) %>% names()
visit_helpers_yesmask_fullset1 <- c(visit_helpers_yesmask_fullset1,missingconeswabdoesntexist_yesmask) # add these helpers back

visit_helpers_yesmask_incompl <- visit_helpers_yesmask %>% subset(!(. %in% visit_helpers_yesmask_fullset1))

cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen <- NA

possiblyincomplete_due_to_saliva_yesmask <- visit_helpers_yesmask_incompl %>% subset(visit_helpers_yesmask_sampletypes[.] %>% sapply(function(x) {all(fullset2yesmask %in% x)})) # which ones are missing saliva
incomplete_dueto_saliva_yesmask <- possiblyincomplete_due_to_saliva_yesmask %>% subset(possiblyincomplete_due_to_saliva_yesmask %>% sapply(function(x) {cov_pcr_sera_yesmask_complete[["pcr_screen"]] %>% filter(helper==x & grepl("saliva",tolower(sample_type))) %>% select(ct_sgene_screen,ct_ngene_screen,ct_orf1ab_screen) %>% unlist() %>% `==`(c(40,40,40)) %>% all()})) # which ones have saliva nondetects at screening
incomplete_yesmask <- visit_helpers_yesmask_incompl %>% subset(!(. %in% incomplete_dueto_saliva_yesmask))
  
cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen[cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper %in% incomplete_dueto_saliva_yesmask] <- T

# impute missing saliva samples
impute_yesmask_helpers <- c(incomplete_dueto_saliva_yesmask)
impute_yesmask_df <- matrix(ncol=ncol(cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]),nrow=length(impute_yesmask_helpers)) %>% as.data.frame(stringsAsFactors=F)
colnames(impute_yesmask_df) <- colnames(cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]])
impute_yesmask_df$helper <- impute_yesmask_helpers
impute_yesmask_df$study_id <- impute_yesmask_df$helper %>% strsplit(" ") %>% lapply(head,1) %>% unlist()
impute_yesmask_df$sample_date <- impute_yesmask_df$helper %>% strsplit(" ") %>% lapply(tail,1) %>% unlist() %>% as.Date(origin="1970-01-01")
impute_yesmask_df$sample_id <- c(1:nrow(impute_yesmask_df)) %>% paste0("imputed_saliva_yesmask_",.)
impute_yesmask_df$exp_name <- "imputed saliva row"
impute_yesmask_df$sample_type <- "Saliva"
impute_yesmask_df$ct_ngene_qpcr <- -1
impute_yesmask_df$ct_sgene_qpcr <- -1
impute_yesmask_df$ct_orfgene_qpcr <- -1
impute_yesmask_df$rxn_quant <- 0
impute_yesmask_df$sample_quant <- 0
impute_yesmask_df$culture_status <- "imputed sample"
impute_yesmask_df$exp_date <- impute_yesmask_df$sample_date
impute_yesmask_df$gii_sessnum <- "not a G-II sample"
impute_yesmask_df$mask_yn <- "not a G-II sample"
impute_yesmask_df$mask_type <- "not a G-II sample"
impute_yesmask_df$cough_count <-  -999
impute_yesmask_df$reserv_time <- -999

impute_yesmask_df$nd_lod_loq <- 0
impute_yesmask_df$cens <-  1 # used to be logical, now numeric; cens 1 means sample_quant is 0
impute_yesmask_df$rxn_quant <- 0
impute_yesmask_df$sample_quant <- 0
impute_yesmask_df$mdfy_sample_quant <- 75
impute_yesmask_df$any_rep_lod1 <- F
impute_yesmask_df$all_rep_lod2 <- F
impute_yesmask_df$all_rep_loq <- F
impute_yesmask_df$av_quant <- 0
impute_yesmask_df$loq <- F
impute_yesmask_df$S.gene.dropout <- T # yes, this is true for the particular study IDs in this list

impute_yesmask_df$nd_lod_loq <- 0 %>% as.factor()
impute_yesmask_df$gii <- 0 %>% as.factor()
impute_yesmask_df$gii_reps <- impute_yesmask_df$helper %>% sapply(function(x) {cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] %>% filter(helper==x) %>% pull(gii_reps) %>% unique()})
impute_yesmask_df$saliva_missing_after_screen <-  T

# bring over 274 2020-10-01 from impute_nomask_df
impute_274_2020_10_01 <- impute_nomask_df %>% filter(helper=="274 2020-10-01")
impute_yesmask_df <- impute_yesmask_df %>% bind_rows(impute_274_2020_10_01)

cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] <- cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] %>% filter((helper %in% visit_helpers_yesmask_fullset1) | (saliva_missing_after_screen) | helper=="274 2020-10-01")
cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] <- cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] %>% bind_rows(impute_yesmask_df)

helpers1 <- cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper %>% unique() %>% sort()
if (all(visit_helpers_yesmask %in% cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper)) {
  print("all yesmask sets are considered quantitated")
} else {
  helpers2 <- visit_helpers_yesmask %>% subset(!(. %in% cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper))
  cat(paste0("The following sets are incomplete:\n",helpers2 %>% paste(collapse="\n"),"\n"))
  for (i in 1:length(helpers2)) {
    j <- visit_helpers_yesmask_sampletypes[helpers2[i]] %>% unlist()
    notj <- fullset1yesmask %>% subset(!(. %in% j))
    cat(paste0(helpers2[i]," has ",j %>% paste(collapse=";"),"; but it's missing ",notj %>% paste(collapse = ",")))
  }
}

# now that cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] has reduced samples, filter demo1 and sero1 to exclude subjects that are missing, and filter pcr_screen to exclude visits that are not in qpcr_allsamples
cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen[is.na(cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$saliva_missing_after_screen)] <- F

cov_pcr_sera_yesmask_complete[["demohist"]] <- cov_pcr_sera_yesmask_complete[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_yesmask_complete[["sero1"]] <- cov_pcr_sera_yesmask_complete[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_yesmask_complete[["pcr_screen"]] <- cov_pcr_sera_yesmask_complete[["pcr_screen"]] %>% filter(helper %in% cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper)
print(paste0("cov_pcr_sera_yesmask_complete qpcr helpers: ",cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_yesmask_complete rows:")
cov_pcr_sera_yesmask_complete %>% sapply(nrow)
```

For dataset 3, take dataset 2 and exclude visits for which any cultures for positive sample_quant have not been run
```{r complete qPCR+culture set for nomask}
cov_pcr_sera_nomask_complete_culture <- cov_pcr_sera_nomask_complete
#qpcr_nomask_culture <- qpcr_nomask
fullset2anomask <- c("G-II Coarse Aerosol","G-II Fine Aerosol","Phone Swab","Midturbinate Swab","G-II Cone Swab")
visit_helpers_notcultured_nomask <- cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]] %>% filter((sample_type %in% fullset2anomask) & culture_status=="not cultured") %>% pull(helper) %>% unique()
cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]] <- cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]] %>% filter(!(helper %in% visit_helpers_notcultured_nomask))
qpcr_nomask_notcultured <- cov_pcr_sera_nomask_complete[["qpcr_allsamples"]] %>% filter(helper %in% visit_helpers_notcultured_nomask)

# which helper groups in nomask (all) have breath samples greater than zero
poshelpers_nomask <- cov_pcr_sera_nomask[["qpcr_allsamples"]] %>% filter(rxn_quant>0 & grepl("G-II",sample_type)) %>% select(helper) %>% unlist() %>% unique() %>% sort()
cat(paste0(length(poshelpers_nomask)," groups that have rxn_quant>0 in G-II types"))
posculturecompletehelpers_nomask <- cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$helper %>% unique() %>% sort() %>% subset(. %in% poshelpers_nomask)
cat(paste0(length(posculturecompletehelpers_nomask)," of these have culture complete"))
poscultureincompletehelpers_nomask <-poshelpers_nomask %>% subset(!(. %in% posculturecompletehelpers_nomask))
cat(paste0(length(poscultureincompletehelpers_nomask)," of these have culture incomplete"))

listofnotcultured_nomask <- cov_pcr_sera_nomask[["qpcr_allsamples"]] %>% filter((helper %in% poscultureincompletehelpers_nomask) & (culture_status=="not cultured") & (sample_type %in% fullset2anomask))
listofnotcultured_nomask <- listofnotcultured_nomask %>% select(study_id,sample_date,sample_type) %>% unique() %>% arrange(study_id,sample_date,sample_type)
write.csv(listofnotcultured_nomask,paste0("incomplete_culture_toship_nomask_",Sys.Date() %>% format("%Y%m%d"),".csv"),row.names=F) # this list does not pick up cone swabs

cov_pcr_sera_nomask_complete_culture[["demohist"]] <- cov_pcr_sera_nomask_complete_culture[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_nomask_complete_culture[["sero1"]] <- cov_pcr_sera_nomask_complete_culture[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_nomask_complete_culture[["pcr_screen"]] <- cov_pcr_sera_nomask_complete_culture[["pcr_screen"]] %>% filter(helper %in% cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$helper)
print(paste0("cov_pcr_sera_nomask_complete_culture qpcr helpers: ",cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_nomask_complete_culture rows:")
cov_pcr_sera_nomask_complete_culture %>% sapply(nrow)
```
culture sets for yes masked samples
```{r complete qPCR+culture set for yesmask}
cov_pcr_sera_yesmask_complete_culture <- cov_pcr_sera_yesmask_complete
#qpcr_yesmask_culture <- qpcr_yesmask
fullset2ayesmask <- c("G-II Coarse Aerosol mask","G-II Fine Aerosol mask","Phone Swab","Midturbinate Swab","G-II Cone Swab mask")
visit_helpers_notcultured_yesmask <- cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]] %>% filter((sample_type %in% fullset2ayesmask) & culture_status=="not cultured") %>% pull(helper) %>% unique()
cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]] <- cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]] %>% filter(!(helper %in% visit_helpers_notcultured_yesmask))
qpcr_yesmask_notcultured <- cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]] %>% filter(helper %in% visit_helpers_notcultured_yesmask)

# which helper groups in yesmask (all) have breath samples greater than zero
poshelpers_yesmask <- cov_pcr_sera_yesmask[["qpcr_allsamples"]] %>% filter(rxn_quant>0 & grepl("G-II",sample_type)) %>% select(helper) %>% unlist() %>% unique() %>% sort()
cat(paste0(length(poshelpers_yesmask)," groups that have rxn_quant>0 in G-II types"))
posculturecompletehelpers_yesmask <- cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$helper %>% unique() %>% sort() %>% subset(. %in% poshelpers_yesmask)
cat(paste0(length(posculturecompletehelpers_yesmask)," of these have culture complete"))
poscultureincompletehelpers_yesmask <-poshelpers_yesmask %>% subset(!(. %in% posculturecompletehelpers_yesmask))
cat(paste0(length(poscultureincompletehelpers_yesmask)," of these have culture incomplete"))

listofnotculturedyesmask <- cov_pcr_sera_yesmask[["qpcr_allsamples"]] %>% filter((helper %in% poscultureincompletehelpers_yesmask) & (culture_status=="not cultured") & (sample_type %in% fullset2ayesmask))
listofnotculturedyesmask <- listofnotculturedyesmask %>% select(study_id,sample_date,sample_type) %>% unique() %>% arrange(study_id,sample_date,sample_type)
write.csv(listofnotculturedyesmask,paste0("incomplete_culture_toship_yesmask_",Sys.Date() %>% format("%Y%m%d"),".csv"),row.names=F)

cov_pcr_sera_yesmask_complete_culture[["demohist"]] <- cov_pcr_sera_yesmask_complete_culture[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_yesmask_complete_culture[["sero1"]] <- cov_pcr_sera_yesmask_complete_culture[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_yesmask_complete_culture[["pcr_screen"]] <- cov_pcr_sera_yesmask_complete_culture[["pcr_screen"]] %>% filter(helper %in% cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$helper)
print(paste0("cov_pcr_sera_yesmask_complete_culture qpcr helpers: ",cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_yesmask_complete_culture rows:")
cov_pcr_sera_yesmask_complete_culture %>% sapply(nrow)
```
create a new set cov_pcr_sera_paired, with complete qpcr and complete culture
```{r paired sets}
paired_helpers <- intersect(visit_helpers_nomask,visit_helpers_yesmask)

# cov_pcr_sera_paired is cov_pcr_sera where helpers are restricted to days where there are masked and unmasked samples
cov_pcr_sera_paired <- cov_pcr_sera
cov_pcr_sera_paired[["qpcr_allsamples"]] <- cov_pcr_sera_paired[["qpcr_allsamples"]] %>% filter(helper %in% paired_helpers)
cov_pcr_sera_paired[["demohist"]] <- cov_pcr_sera_paired[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_paired[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired[["sero1"]] <- cov_pcr_sera_paired[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_paired[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired[["pcr_screen"]] <- cov_pcr_sera_paired[["pcr_screen"]] %>% filter(study_id %in% cov_pcr_sera_paired[["qpcr_allsamples"]]$study_id)
print(paste0("cov_pcr_sera_paired qpcr helpers: ",cov_pcr_sera_paired[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_paired rows:")
cov_pcr_sera_paired %>% sapply(nrow)

# cov_pcr_sera_paired_complete is cov_pcr_sera_paired restricted to person-days with complete qpcr sets for both masked and unmasked sets

paired_helpers_fullqpcr <- intersect(cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper,cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper) %>% unique()

cov_pcr_sera_paired_complete <- cov_pcr_sera_paired
cov_pcr_sera_paired_complete[["qpcr_allsamples"]] <- cov_pcr_sera_paired_complete[["qpcr_allsamples"]] %>% filter(helper %in% paired_helpers_fullqpcr)

impute_paired_df <- bind_rows(impute_nomask_df,impute_yesmask_df) %>% distinct(helper,.keep_all = T) # imputation for the yesmask and nomask sets was duplicated, as they're salivas

cov_pcr_sera_paired_complete[["qpcr_allsamples"]] <- cov_pcr_sera_paired_complete[["qpcr_allsamples"]] %>% bind_rows(impute_paired_df)

cov_pcr_sera_paired_complete[["demohist"]] <- cov_pcr_sera_paired_complete[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired_complete[["sero1"]] <- cov_pcr_sera_paired_complete[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired_complete[["pcr_screen"]] <- cov_pcr_sera_paired_complete[["pcr_screen"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete[["qpcr_allsamples"]]$study_id)
print(paste0("cov_pcr_sera_paired_complete qpcr helpers: ",cov_pcr_sera_paired_complete[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_paired_complete rows:")
cov_pcr_sera_paired_complete %>% sapply(nrow)

# cov_pcr_sera_paired_complete_culture
cov_pcr_sera_paired_complete_culture <- cov_pcr_sera_paired_complete
fullsampletypesetforpairedculture <- c("Phone Swab","Midturbinate Swab",
                                       "G-II Coarse Aerosol","G-II Fine Aerosol","G-II Cone Swab",
                                       "G-II Coarse Aerosol mask","G-II Fine Aerosol mask","G-II Cone Swab mask")
visit_helpers_notcultured_paired <- cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]] %>% filter((sample_type %in% fullsampletypesetforpairedculture) & culture_status=="not cultured") %>% pull(helper) %>% unique()
cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]] <- cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]] %>% filter(!(helper %in% visit_helpers_notcultured_paired))
qpcr_paired_notcultured <- cov_pcr_sera_paired_complete[["qpcr_allsamples"]] %>% filter(helper %in% visit_helpers_notcultured_paired)

cov_pcr_sera_paired_complete_culture[["demohist"]] <- cov_pcr_sera_paired_complete_culture[["demohist"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired_complete_culture[["sero1"]] <- cov_pcr_sera_paired_complete_culture[["sero1"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]]$study_id)
cov_pcr_sera_paired_complete_culture[["pcr_screen"]] <- cov_pcr_sera_paired_complete_culture[["pcr_screen"]] %>% filter(study_id %in% cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]]$study_id)
print(paste0("cov_pcr_sera_paired_complete_culture qpcr helpers: ",cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]]$helper %>% unique() %>% length()))
print("cov_pcr_sera_paired_complete_culture rows:")
cov_pcr_sera_paired_complete_culture %>% sapply(nrow)
```
clean up and write
```{r write files}
cov_pcr_sera_nomask[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_nomask[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_nomask_complete[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_nomask_complete[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_nomask_complete_culture[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_nomask_complete_culture[["pcr_screen"]]$helper <- NULL

cov_pcr_sera_yesmask[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_yesmask[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_yesmask_complete[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_yesmask_complete[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_yesmask_complete_culture[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_yesmask_complete_culture[["pcr_screen"]]$helper <- NULL

cov_pcr_sera_paired[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_paired[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_paired_complete[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_paired_complete[["pcr_screen"]]$helper <- NULL
cov_pcr_sera_paired_complete_culture[["qpcr_allsamples"]]$helper <- NULL
cov_pcr_sera_paired_complete_culture[["pcr_screen"]]$helper <- NULL

nomask <- list(cov_pcr_sera_nomask,cov_pcr_sera_nomask_complete,cov_pcr_sera_nomask_complete_culture)
names(nomask) <- c("cov_pcr_sera_nomask","cov_pcr_sera_nomask_complete","cov_pcr_sera_nomask_complete_culture")
yesmask <- list(cov_pcr_sera_yesmask,cov_pcr_sera_yesmask_complete,cov_pcr_sera_yesmask_complete_culture)
names(yesmask) <- c("cov_pcr_sera_yesmask","cov_pcr_sera_yesmask_complete","cov_pcr_sera_yesmask_complete_culture")
paired <- list(cov_pcr_sera_paired,cov_pcr_sera_paired_complete,cov_pcr_sera_paired_complete_culture)
names(paired) <- c("cov_pcr_sera_paired","cov_pcr_sera_paired_complete","cov_pcr_sera_paired_complete_culture")

objectstowrite <- "cov_pcr_sera_" %>% paste0(c("nomask","yesmask","paired")) %>% lapply(paste0,c("","_complete","_complete_culture")) %>% unlist()
filenamestowrite <- "cov_pcr_sera_" %>% paste0(Sys.Date() %>% format("%Y%m%d"),"_") %>% paste0(c("nomask","yesmask","paired")) %>% lapply(paste0,c("","_complete","_complete_culture")) %>% unlist() %>% paste0(".RDS")

for (i in 1:length(objectstowrite)) {
  j <- eval(parse(text=objectstowrite[i]))
  saveRDS(j,file = filenamestowrite[i])
}
```

