---
title: "Initial merging of all files"
author: Jake & Sanmi, modified by Filbert
date: "1/22/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(readxl)
```
Source data from
1) qPCR results
2) screening PCR results partially merged with temp & sx_current
3) culture results latest
4) serology results latest
5) StopCOVID REDCap projects, deidentifed

Output:
1) demohist table, demographics and medical history and all other variables which exist only once for each subject ID
2) screening PCR table, with merged data pertaining to clinical visits
3) qPCR table, merged with culture data
4) serology table
Bundled in one R object that is a list() of all four tables
```{r read source files}
sx <- read.csv(file="../source_data/ncovPCRtempsxresults1_210805.csv",stringsAsFactors = F) # screening PCR data, merged with temperature and symptoms reported that day (if available)
rc <- readRDS("../source_data/ncovrcmain-deid.RData") # this is the main REDCap project that the clinic interacts with
rcdt <- readRDS("../source_data/ncovrcdt-deid.RData") # this REDCap project has the surveys sent to participants on a daily basis via text message

nrow(sx)
head(sx)
sx %>% group_by(redcap_event_name) %>% summarise(n=n())

nrow(rc)
head(rc)

nrow(rcdt)
head(rcdt)
```
Onset of symptoms dates from rc:
In the baseline survey, participants are asked when they thought their symptoms started. If they came in for shedding or contact visits, participants who reported symptoms were also asked (again, sometimes multiple times) when their symptoms were started. Occasionally this leads to conflicting reports in the date. Generally we took the earliest date as the most accurate, but for a couple participants we felt their most accurate answers came from email communications (before they had enrolled)
```{r onset sx calcs}
covidsubjs <- rc %>% filter(covid_pos=='1') %>% pull(subject_id) %>% as.numeric() %>% sort() # covid_pos is set to 1 if the screening PCR was positive
allsymptoms <- c("nose_run","nose_stuf","sneeze","throat_sr","earache","malaise","headache","mj_ache","sw_fever_chill","nausea","loa","vomit","diarrhea","chest_tight","sob","cough","taste_smell_yn")

# primary source of onset of symptoms date comes from "Symptom Card Retrospective"
onsetdates <- rc %>% filter((subject_id %in% covidsubjs) & !is.na(time_feel_sick_ret)) %>% select(subject_id,time_feel_sick_ret)
onsetdates$onsetdate_ret <- onsetdates$time_feel_sick_ret %>% substr(1,10) %>% as.Date(origin="1970-01-01") %>% as.character()

# we collected onset of symptoms date from the daily symptom card too
sxdates <- rc %>% filter((subject_id %in% covidsubjs) & !is.na(symptom_time) & (grepl("shedding|contact",redcap_event_name))) %>% select(subject_id,symptom_time,redcap_event_name,sx_days)
sxdates$symptom_time <- sxdates$symptom_time %>% substr(1,10) %>% as.Date(origin="1970-01-01")
sxdates$calculated_onset <- (sxdates$symptom_time - as.numeric(sxdates$sx_days)) %>% as.character()
collapsed_onsetdates <- covidsubjs %>% sapply(function(x) {sxdates %>% filter(subject_id==x) %>% pull(calculated_onset) %>% na.omit() %>% sort() %>% unique() %>% paste(collapse=" ")}) # collapsed onset dates into a single text field

sxdates_collated <- c(covidsubjs,collapsed_onsetdates) %>% matrix(ncol=2) %>% as.data.frame(stringsAsFactors=F) # make it into a data frame for convenience
colnames(sxdates_collated) <- c("subject_id","onset_from_sx_card")
onsetdates <- onsetdates %>% full_join(sxdates_collated,by="subject_id") # this puts together
onsetdates <- onsetdates %>% arrange(subject_id) %>% select(-c("time_feel_sick_ret"))
onsetdates[is.na(onsetdates)] <- ""
onsetdates$subject_id <- onsetdates$subject_id %>% as.numeric()

# decide to take the earliest onset date as the "real" onset date
onsetdates$real_onset <- paste(onsetdates$onsetdate_ret,onsetdates$onset_from_sx_card) %>% 
  lapply(function(x) {
    if (x==" ") {
      NA
    } else {
      x %>% strsplit(" ") %>% unlist() %>% subset(.!="") %>% as.Date(origin="1970-01-01") %>% min() # if there are multiple possible dates, take the earliest date
    }
  }) %>% unlist() %>% as.Date(origin="1970-01-01")
summary(onsetdates$real_onset)

# there are 6 NAs because these people had neither onset_from_sx_card or onsetdate_ret
# 323 had onsetdate_ret but not onset_from_sx_card. That's because 323 declined to report the number of days from onset on 11/4, and then had no symptoms on 11/6.

# corrections curated by Sanmi for subjects 88 and 336
# 336 (11/11/2020) and 88 (7/11/2020)
onsetdates<-rbind(onsetdates,onsetdates[nrow(onsetdates)+1,])
onsetdates$subject_id[nrow(onsetdates)]<-88
onsetdates$real_onset[nrow(onsetdates)]<-"2020-07-11"
onsetdates<-rbind(onsetdates,onsetdates[nrow(onsetdates)+1,])
onsetdates$subject_id[nrow(onsetdates)]<-336
onsetdates$real_onset[nrow(onsetdates)]<-"2020-11-11"

nrow(onsetdates)
head(onsetdates) # note that some participants will have no onset dates if they didn't have symptoms
summary(onsetdates)
```
demohist: demographics, medical history, things that depend only on subject_id and not on date
```{r demographics 1}
# Age: NA to fill in outliers i.e. those entering today's date and birthdate
rc[which(rc$calc_eligibility<1),"subject_id"] # subject ids of wrongly entered date of births
hist(as.numeric(rc$calc_eligibility),xlab = "Age")
rc[which(rc$calc_eligibility<1),"calc_eligibility"] <- NA # set them to NA

# the age variable is in the subject tracker event; other variables are in baseline survey
demohist_base <- rc %>% filter(grepl("baseline",redcap_event_name))
nrow(demohist_base) # which equals the number of records created, but not all records turned out to be enrolled participants
subj_age <- rc %>% filter(grepl("subject",redcap_event_name)) %>% select(subject_id,calc_eligibility)
demohist_base$calc_eligibility <- NULL # we're replacing this column
demohist_base <- demohist_base %>% left_join(subj_age,by="subject_id")

# remove participants who did not consent and finish the baseline survey
demohist_enrolled <- demohist_base %>% filter(survey_conclusion_complete!=0)

# demographics race questions
racecols <- colnames(rc) %>% subset(grepl("race_",.)) 
# chronic respiratory illness, drinks per week, race, BMI, ox_sat, symptoms_ret
demohistcols <- c("subject_id","redcap_event_name","calc_eligibility","sex_birth",
              "asthma","allerg","copd","cf","bronchitis","emphys","lung_other",
              "alc_week","drinks_p_day",
              "cigarette_ever",
              "singing_lessons","singing_ensemble",
              "latino",
              racecols, "other_race")
demohist_all <- demohist_enrolled %>% select(all_of(demohistcols))

# "tag" variable just keeps track of which arm a participant entered into the study
demohist_all$tag<- demohist_all$redcap_event_name %>% gsub("\\D+","",.)
demohist_all$tag[demohist_all$redcap_event_name=="baseline_arm_1"]<-"Surveillance entry"
demohist_all$tag[demohist_all$redcap_event_name=="case_baseline_arm_2"]<-"Case entry"
demohist_all$tag[demohist_all$redcap_event_name=="contact_baseline_arm_3"]<-"Contact entry"
demohist_all$tag[demohist_all$redcap_event_name=="baseline_followup_arm_4"]<-"IR entry"
demohist_all$redcap_event_name <- NULL

# chronic respiratory illness
cri_col <-c("asthma","allerg","copd","cf","bronchitis","emphys","lung_other")
demohist_all$cri <- demohist_all[,cri_col] %>% apply(1,function(x) {if (any(x==1)) {1} else {0}}) 
demohist_all[,cri_col] <- NULL # cri columns are temporary

# drinks per week
demohist_all$drinks_p_week <- as.numeric(demohist_all$drinks_p_day) * as.numeric(demohist_all$alc_week)
demohist_all$drinks_p_week[which(is.na(demohist_all$drinks_p_week))] <- 0
demohist_all$drinks_p_week[which(as.numeric(demohist_all$alc_week)>7)] <- NA # someone put drinks_p_week = 41

# racelabels
racelabels<-c("White","Black or African American","American Indian or Alaska Native","Asian Indian","Chinese","Filipino","Japanese","Korean","Vietnamese","Cambodian","Laotian","Other Asian","Native Hawaiian","Other Pacific Islander",
"Other race","Don't Know","Decline")
racecolswhich <- which(colnames(demohist_all) %in% racecols)
colnames(demohist_all)[racecolswhich] <- racelabels

# most of these "other" races are not actually races
otherraceanswers <- c("Mexican","Hispanic","Jewish/Iranian","Jewish","Middle Eastern","Lebanese")
demohist_all$`Other race`[demohist_all$other_race %in% otherraceanswers] <- "0"
demohist_all$White[demohist_all$other_race %in% otherraceanswers] <- "1"
demohist_all$race <- demohist_all[,racelabels] %>% apply(1,function(x) {x %>% subset(.==1) %>% names() %>% paste(collapse="-")})

# check that all participants checked at least one race box off
demohist_enrolled %>% select(racecols) %>% apply(1,function(x) {x %>% as.numeric() %>% sum()}) %>% unique()

demohist_all$multiracial <- demohist_all$race %>% sapply(function(x) {if (grepl("-",x)) {"Yes"} else {"No"}})
demohist_all[,racelabels] <- NULL #remove race columns

# BMI data frame
bmi <- rc %>% filter(!is.na(bmi)) %>% select(subject_id,redcap_event_name,bmi)
# 255 provided bmi twice, due to becoming a case twice. Use the average
bmidupesubjs <- bmi$subject_id %>% subset(duplicated(.)) %>% unique()
for (i in bmidupesubjs) {
  j <- bmi %>% filter(subject_id==i)
  k <- j[1,]
  k$bmi[1] <- j$bmi %>% as.numeric() %>% mean() %>% round(2)
  bmi <- bmi %>% filter(subject_id!=i)
  bmi <- bmi %>% bind_rows(k)
}

# define symptom categories: upper respiratory, lower respiratory, systemic, and gi

ur_sx <- c("nose_run","nose_stuf", "sneeze", "throat_sr", "earache")
lr_sx <- c("chest_tight", "sob", "cough")
sys_sx <- c("malaise", "headache", "mj_ache", "sw_fever_chill")
gi_sx <- c("loa", "vomit", "diarrhea", "nausea")

ur_sx_ret <- ur_sx %>% paste0("_ret")
lr_sx_ret <- lr_sx %>% paste0("_ret")
sys_sx_ret <- sys_sx %>% paste0("_ret")
gi_sx_ret <- gi_sx %>% paste0("_ret")
all_sx_ret <- c(ur_sx_ret,lr_sx_ret,sys_sx_ret,gi_sx_ret)
smell_medcare <- c("taste_smell_yn_covid_ret","sx_med_care_ret")

# retrospective symptoms at peak severity was asked at IR follow-up, so the baseline ret values are excluded
ret_symp <- rc %>% filter(grepl("ir",redcap_event_name)) %>%
  select(subject_id,all_of(all_sx_ret),all_of(smell_medcare))

# there was someone who filled out two IR surveys - they had scheduled a visit then postponed it - so need to dedupe ret_symp 
retsympdupesubjs <- ret_symp$subject_id %>% subset(duplicated(.)) %>% unique()
for (i in retsympdupesubjs) {
  j <- ret_symp %>% filter(subject_id==i)
  k <- j[1,]
  k[1,all_sx_ret] <- j[,all_sx_ret] %>% apply(2,max)
  ret_symp <- ret_symp %>% filter(subject_id!=i)
  ret_symp <- ret_symp %>% bind_rows(k)
}

ret_symp[,all_sx_ret] <- ret_symp[,all_sx_ret] %>% apply(2,as.numeric)
ret_symp$ur_ret <- ret_symp[,ur_sx_ret] %>% apply(1,sum)
ret_symp$lr_ret <- ret_symp[,lr_sx_ret] %>% apply(1,sum)
ret_symp$sys_ret <- ret_symp[,sys_sx_ret] %>% apply(1,sum)
ret_symp$gi_ret <- ret_symp[,gi_sx_ret] %>% apply(1,sum)

ret_symp %>% filter(is.na(ur_ret)) %>% select(all_of(ur_sx_ret)) %>% unlist() %>% unique()
ret_symp %>% filter(is.na(lr_ret)) %>% select(all_of(lr_sx_ret)) %>% unlist() %>% unique()
ret_symp %>% filter(is.na(sys_ret)) %>% select(all_of(sys_sx_ret)) %>% unlist() %>% unique()
ret_symp %>% filter(is.na(gi_ret)) %>% select(all_of(gi_sx_ret)) %>% unlist() %>% unique()
# checked that ret sum scores are NA because all ret symptoms are NA

# merge demohist_all, ret_symp, bmi, onset dates by subject id

demohist <- demohist_all %>% left_join(bmi %>% select(subject_id,bmi),by="subject_id") # attach BMI
demohist <- demohist %>% left_join(ret_symp,by="subject_id") # attach restrospective symptoms
onsetdates$subject_id <- onsetdates$subject_id %>% as.character()
demohist <- demohist %>% left_join(onsetdates %>% select(subject_id,real_onset),by="subject_id") # attach symptom onset dates
# demohist$survey_date <- NULL
demohist <- demohist %>% rename(age_entry=calc_eligibility)
demohist$age_entry <- demohist$age_entry %>% as.numeric() %>% floor()
demohist <- demohist %>% rename(study_id=subject_id)

# define the date_of_enrollment as the date of a participant's first clinic visit
demohist$date_of_enrollment <- demohist$study_id %>%
  lapply(function(x) {rc %>% filter(subject_id==x & !is.na(time_of_visit)) %>% pull(time_of_visit) %>% substr(1,10)}) %>%
  lapply(function(x) {if (length(x)==0) {NA} else {x}}) %>%
  lapply(function(x) {x %>% as.Date(origin="1970-01-01") %>% min()}) %>%
  unlist() %>%
  as.Date(origin="1970-01-01")

demohist$date_of_enrollment %>% unique()
demohist$date_of_enrollment %>% summary()

nrow(demohist)
head(demohist)
summary(demohist)
```
screening pcr data frame

```{r screening pcr}
# a pulse oximeter was used to measure SPO2 at every visit
spo2 <- rc %>% filter(!is.na(ox_sat)) %>% select(subject_id,redcap_event_name,time_pulseox_measured,ox_sat)
spo2$date_pulseox_measured <- spo2$time_pulseox_measured %>% substr(1,10) %>% as.Date(origin="1970-01-01")

# merge spo2 to sx (pcr results by subject id and sample date)
sx$helper1 <- paste(sx$subject_id,sx$sample_date)
spo2$helper1 <- paste(spo2$subject_id,spo2$date_pulseox_measured)
pcr_sx_temp_spo2 <- sx %>% left_join(spo2 %>% select(helper1,ox_sat),by="helper1")
pcr_sx_temp_spo2$helper1 <- NULL

# oral_temp, fever_samp, fever_ever (modify demohist), days_past_onset = as.numeric(date_visit - real_onset)
pcr_screen <- pcr_sx_temp_spo2
pcr_screen <- pcr_screen %>% rename(study_id=subject_id)
pcr_screen$study_id <- pcr_screen$study_id %>% as.character()

pcr_screen$oral_temp <- pcr_screen$oral_temp %>% sapply(function(x) {if (!is.na(x)) {if (x>50) {((x- 32 ) * 5/9) %>% round(2)} else {x}} else {x}}) # if !NA, then check if temperatures were recorded in F instead of C
pcr_screen$fever_samp <- (pcr_screen$oral_temp>=37.8) %>% as.numeric() # 37.8 is our definition of a fever
pcr_screen$sample_date <- pcr_screen$sample_date %>% as.Date(origin="1970-01-01")
pcr_screen <- pcr_screen %>% left_join(demohist %>% select(study_id,real_onset),by="study_id")
pcr_screen$days_past_onset <- as.numeric(pcr_screen$sample_date-pcr_screen$real_onset)
pcr_screen$real_onset <- NULL

pcr_screen$ct_avg_screen <- pcr_screen[,c("S.gene","N.gene","ORF1ab")] %>% apply(1,mean)
pcr_screen <- pcr_screen %>% rename(ct_sgene_screen=S.gene,ct_ngene_screen=N.gene,ct_orf1ab_screen=ORF1ab)
pcr_screen <- pcr_screen %>% rename(tac_id=TAC.ID)
pcr_screen$helper2 <- NULL

# attach compound symptom scores
pcr_screen$ur <- pcr_screen[,ur_sx] %>% apply(1,sum)
pcr_screen$lr <- pcr_screen[,lr_sx] %>% apply(1,sum)
pcr_screen$gi <- pcr_screen[,gi_sx] %>% apply(1,sum)
pcr_screen$sys <- pcr_screen[,sys_sx] %>% apply(1,sum)

summary(pcr_screen)

# attach fever_ever to demohist
#demohist$fever_ever <- demohist$study_id %>% sapply(function(x) {pcr_screen %>% filter(study_id==x) %>% pull(fever_samp) %>% sum() %>% `>`(0)})
fever_ever <- demohist$study_id %>% sapply(function(x) {pcr_screen %>% filter(study_id==x) %>% pull(fever_samp) %>% na.omit()})
fever_ever <- fever_ever %>% lapply(function(x) {if (length(x)==0) {NA} else {x}}) # convert character(0) to NA
fever_ever <- fever_ever %>% lapply(function(x) {x %>% sum() %>% `>`(0)}) %>% unlist()

demohist$fever_ever <- fever_ever
summary(demohist$fever_ever)
```
quantitative PCR

Conversion of reaction quantity (rxn_quant) to sample quantity (sample_quant), in RNA copies per mL:
Multiply rxn_quant by 20.8333 (=125/6) for saliva samples, by 25 for all other samples. These were the standard dilution from extracted samples.

"There are a couple of samples (NA0035966889 & NA0035967370) that need a different ratio because they didn’t have enough volume for extraction." - hardcoded below as 125/3 and 39.0635 respectively
```{r quantitative PCR data}
qpcrnewfile <-"../source_data/RT-qPCR_results_20210611.csv"
qpcrnew <- read.csv(file=qpcrnewfile,stringsAsFactors=F)

# get subject id, date of collection, and sample type from StopCOVID sample listing
allsamples <- read.csv("../source_data/StopCOVIDsamples.csv",colClasses = "character") %>% select(study_id, sample_type, sample_id, date_collected_sg,event_id) %>% rename(sample_date=date_collected_sg,redcap_event_name=event_id)

nrow(allsamples)

# samples up to what date
max(qpcrnew$exp_date) %>% substr(1,8) %>% as.Date(origin="1970-01-01",format="%Y%m%d")

# merge pcr with StopCOVID allsamples list
qpcr_allsamples <- qpcrnew %>% left_join(allsamples %>% select(study_id,sample_type,sample_id,sample_date), by = "sample_id")
nrow(qpcr_allsamples)

qpcr_allsamples$sample_quant <- qpcr_allsamples$rxn_quant * 25
qpcr_allsamples$sample_quant[qpcr_allsamples$sample_type=="Saliva Derivative"] <- qpcr_allsamples$rxn_quant[qpcr_allsamples$sample_type=="Saliva Derivative"] * (125/6)
qpcr_allsamples$sample_quant[qpcr_allsamples$sample_id=="NA0035967370"] <- qpcr_allsamples$rxn_quant[qpcr_allsamples$sample_id=="NA0035967370"] * 39.0635
qpcr_allsamples$sample_quant[qpcr_allsamples$sample_id=="NA0035966889"] <- qpcr_allsamples$rxn_quant[qpcr_allsamples$sample_id=="NA0035966889"] * (125/3)
qpcr_allsamples <- qpcr_allsamples %>% rename(exp_name=exp_date)
qpcr_allsamples$exp_date <- qpcr_allsamples$exp_name %>% substr(1,8) %>% as.Date(origin="1970-01-01",format="%Y%m%d")

nrow(qpcr_allsamples)
summary(qpcr_allsamples)
```
# Read in the mask, cough, G-II session number, G-II reservoir time data and merge with the qRT-PCR plus sample identify key dataset; clean
cough_count set to -1 if for some reason the operator left the column blank.
cough_count and reserv_time are set to -999 if not a G-II sample.
```{r mask cough G-II data}
masks_cough <- read.csv("../source_data/ncovg2samplesmasksyn.csv",colClasses = "character")
nrow(masks_cough)
head(masks_cough)

masks_cough_clean1 <- masks_cough
masks_cough_clean1$mask_type[is.na(masks_cough_clean1$mask_type)] <- "no mask"
masks_cough_clean1$cough_count[is.na(masks_cough_clean1$cough_count)] <- 0 # the missing ones were in fact done, no coughs, operators forgot to record
masks_cough_clean1$cone_barcode[is.na(masks_cough_clean1$cone_barcode)] <- ""
allbarcodefields <- colnames(masks_cough_clean1) %>% subset(grepl("fine|coarse|cone",.))
masks_cough_clean1$allbarcodes <- masks_cough_clean1[,allbarcodefields] %>% apply(1,function(x) {x %>% unlist() %>% paste(collapse=";")})

# transfer mask, cough, reservoir time, and session number over to qpcr_allsamples
maskscoughtransfercols <- c("sessnum","mask_yn","mask_type","cough_count","reserv_time")
maskstransfercolsfac <- c("sessnum","mask_type","mask_yn") 
maskstransfercolsnum <- c("cough_count","reserv_time")
for (i in 1:nrow(qpcr_allsamples)) {
  if (grepl("G2|G-II",qpcr_allsamples$sample_type[i])) {
    j <- masks_cough_clean1 %>% filter(grepl(qpcr_allsamples$sample_id[i],allbarcodes))
    qpcr_allsamples[i,maskscoughtransfercols] <- j[1,maskscoughtransfercols]
  } else {
    qpcr_allsamples[i,maskstransfercolsfac] <- "not a G-II sample"
    qpcr_allsamples[i,maskstransfercolsnum] <- "-999" # not a G-II sample but numeric
  }
}
qpcr_allsamples <- qpcr_allsamples %>% rename(gii_sessnum=sessnum)
summary(qpcr_allsamples)
```
Additional breath sample data cleaning
- Adjust for reservoir time: adjust sample_quant and cough counts to account for time if less than 30 min -- standardize these variables as per 30 min in order to allow for like-comparisons between samples.
- Give the masked G-II sample replicates an indicator as part of their name that they are masked samples
- Add a cens variable indicator if replicate is nondetect
- Add an indicator if replicate is nondetectable (=0), between LOD and LOQ (=1), or above LOQ (=2)
- Add an indicator if G-II samples present for study_id-date_visit combination
```{r additional breath sample data cleaning}
# attach "mask" to sample type if mask_yn=="1"
qpcr_allsamples$sample_type[qpcr_allsamples$mask_yn=="1"] <- qpcr_allsamples$sample_type[qpcr_allsamples$mask_yn=="1"] %>% paste("mask")

# old variable cens sample_quant<250
# qpcr_allsamples$cens <- (qpcr_allsamples$sample_quant<250) %>% as.numeric()
# new variable cens
qpcr_allsamples$cens <- (qpcr_allsamples$sample_quant==0) %>% as.numeric()

# new variable nd_lod_loq is 0 if sample_quant=0, 1 if sample_quant<250, nd_lod_loq is 2 if sample_quant>=250
qpcr_allsamples$nd_lod_loq[qpcr_allsamples$sample_quant==0] <- 0
qpcr_allsamples$nd_lod_loq[which(qpcr_allsamples$sample_quant<250 & qpcr_allsamples$sample_quant>0)] <- 1
qpcr_allsamples$nd_lod_loq[qpcr_allsamples$sample_quant>=250] <- 2

# gii indicates if a gii session
# gii_reps is the number of gii replicates = sum(gii)
qpcr_allsamples$gii <- (qpcr_allsamples$gii_sessnum!="not a G-II sample") %>% as.numeric()
qpcr_allsamples <- qpcr_allsamples %>% group_by(study_id,sample_date) %>% mutate(gii_reps=sum(gii)) %>% ungroup()

# remove "Derivative"
qpcr_allsamples$sample_type <- qpcr_allsamples$sample_type %>% gsub(" Derivative","",.)

# adjust sample_quant and cough: if reserve_time<30, sample_quant and cough should be scaled proportionately
for (i in 1:nrow(qpcr_allsamples)) {
  if (qpcr_allsamples$reserv_time[i]!=-999) {
    if (qpcr_allsamples$reserv_time[i]<30) {
      qpcr_allsamples$sample_quant[i] <- qpcr_allsamples$sample_quant[i]*(30/as.numeric(qpcr_allsamples$reserv_time[i]))
      qpcr_allsamples$cough_count[i] <- as.numeric(qpcr_allsamples$cough_count[i])*(30/as.numeric(qpcr_allsamples$reserv_time[i]))
    }
  }
}

# additional lod loq variable indicators (originally in comb_then_split)

qpcr_allsamples$any_rep_lod1 <- qpcr_allsamples$sample_id %>% sapply(function(x) {qpcr_allsamples %>% filter(sample_id==x) %>% pull(nd_lod_loq) %>% `>`(0) %>% any()})

qpcr_allsamples$all_rep_lod2 <- qpcr_allsamples$sample_id %>% sapply(function(x) {qpcr_allsamples %>% filter(sample_id==x) %>% pull(nd_lod_loq) %>% `>`(0) %>% all()})

qpcr_allsamples$all_rep_loq <- qpcr_allsamples$sample_id %>% sapply(function(x) {qpcr_allsamples %>% filter(sample_id==x) %>% pull(nd_lod_loq) %>% `>`(1) %>% all()})
  
# av_quant (originally in comb_then_split)
qpcr_allsamples$av_quant <- qpcr_allsamples$sample_id %>% sapply(function(x) {qpcr_allsamples %>% filter(sample_id==x) %>% pull(sample_quant) %>% mean()})

# s_gene_pos_id (originally in comb_then_split)
qpcr_allsamples$S.gene.dropout <- qpcr_allsamples$study_id %>% sapply(function(x) {qpcr_allsamples %>% filter(study_id==x) %>% pull(ct_sgene_qpcr) %>% `<`(0) %>% all()})

# loq (originally in comb_then_split)
qpcr_allsamples$loq <- (qpcr_allsamples$av_quant >= 250)

# mdfy_sample_quant (originally in other scripts) 75 as the censored upper limit
qpcr_allsamples$mdfy_sample_quant <- qpcr_allsamples$sample_quant %>% sapply(function(x) {if (x==0) {75} else {x}}) 
```
# fix sample type names
```{r}
pcr_screen$sample_type[pcr_screen$sample_type=="MT"] <- "Midturbinate Swab"
pcr_screen$sample_type[pcr_screen$sample_type=="Anterior Nasal Swab Derivative"] <- "Anterior Nasal Swab"
pcr_screen$sample_type[pcr_screen$sample_type=="Saliva-Pooled"] <- "Saliva"
qpcr_allsamples$sample_type <- qpcr_allsamples$sample_type %>% gsub("G2","G-II",.)
summary(pcr_screen)
summary(qpcr_allsamples)
```
serology data frame
first process the vaccination data from rc and rcdt: when did a participant report covid-19 vaccination?
We only asked which month, but because of the way the surveys were sent and the questions structured, we can narrow it down to a possible window of dates
```{r vaccination1 table}
add.months <- function (date,n) {
  return(seq(date, by = paste (n, "months"), length = 2)[2])
}
maincovidvaccols <- c("vacc_time","covid_shot_yn","covid_shot_m","covid_shot_where","covid_shot_type","covid_second_shot_yn","vacc_time_recent","covid_shot_yn_recent","covid_shot_m_recent","covid_shot_where_recent","covid_shot_type_recent","covid_second_shot_yn_recent","covid_shot2_2weeks","covid_shot2_2weeks_recent")
maincovidshotcols <- c("covid_shot_yn","covid_second_shot_yn","covid_shot_yn_recent","covid_second_shot_yn_recent") # these are just the shot fields

dtcovidcols <- c("time_daily_survey","covid_vacc_yn","covid_vacc_first_m","covid_vacc_sec_yn")
dtcovidshotcols <- c("covid_vacc_yn","covid_vacc_sec_yn")# just the shot fields

vaccmain <- rc %>% filter_at(vars(all_of(maincovidshotcols)),any_vars(!is.na(.))) %>% select(subject_id,redcap_event_name,redcap_repeat_instance,all_of(maincovidvaccols))
vaccdt <- rcdt %>% filter_at(vars(all_of(dtcovidshotcols)),any_vars(!is.na(.))) %>% select(subject_id,redcap_event_name,redcap_repeat_instance,all_of(dtcovidcols))

# who has no vaccine information because of no rows in vaccmain and vaccdt
alldemohistsubjs <- demohist$study_id %>% unique()
novaccinfosubjs <- alldemohistsubjs %>% subset(!(. %in% c(vaccmain$subject_id,vaccdt$subject_id)))

# global first shot vaccination status
vaccinfosubjs <- alldemohistsubjs %>% subset(!(. %in% novaccinfosubjs))

globalvacc1status <- vaccinfosubjs %>% lapply(function(x) {c(
  vaccmain %>% filter(subject_id==x) %>% select(covid_shot_yn,covid_shot_yn_recent) %>% unlist(),
  vaccdt %>% filter(subject_id==x) %>% select(covid_vacc_yn) %>% unlist()
  ) %>% paste(collapse=";")}) # first shot info from both data frames, into a single string with 1s, 0s, NAs
names(globalvacc1status) <- vaccinfosubjs

subjsnotvaccinatedstatus <- globalvacc1status %>% subset(lapply(.,function(x) {!grepl("1",x)}) %>% unlist()) # includes people who answered 9 (decline)
subjsnotvaccinated <- names(subjsnotvaccinatedstatus)

subjsvaccinatedstatus <- globalvacc1status %>% subset(lapply(.,function(x) {grepl("1",x)}) %>% unlist())
subjsvaccinated <- names(subjsvaccinatedstatus)

# demohist$vacc_status is "no info","not vaccinated","got first shot"
demohist$vacc_status <- NA
demohist$vacc_status[which(demohist$study_id %in% novaccinfosubjs)] <- "no info"
demohist$vacc_status[which(demohist$study_id %in% subjsnotvaccinated)] <- "not vaccinated"
demohist$vacc_status[which(demohist$study_id %in% subjsvaccinated)] <- "got first shot"
summary(demohist$vacc_status %>% as.factor())

# figure vaccine reported dates from the daily survey
vacc1dtsubjs <- vaccdt %>% filter(covid_vacc_yn==1) %>% select(subject_id) %>% unlist()
vacc1dthist <- vaccdt %>% filter((subject_id %in% vacc1dtsubjs) & !is.na(covid_vacc_yn))
vaccdtdates <- data.frame()
for (i in vacc1dtsubjs) {
  j <- vacc1dthist %>% filter(subject_id==i)
  j$time_daily_survey <- j$time_daily_survey %>% substr(1,10) %>% as.Date(origin="1970-01-01")
  if (nrow(j)>1) {
    # reduce to just two rows: the date when they reported yes, and the most recent immediately preceding time that they reported no. That gives us the window of vaccination
    j1 <- j %>% arrange(desc(time_daily_survey)) %>% `[`(c(1,2),) %>% arrange(time_daily_survey)
    kearly <- j1$time_daily_survey[1]
    klate <- j1$time_daily_survey[2]
    k <- c(i,as.character(kearly),as.character(klate))
    # check the reported month just in case
    jvaccmonth <- j1$covid_vacc_first_m[2] %>% as.numeric()
    jsurveymonth <- j1$time_daily_survey[2] %>% substr(6,7) %>% as.numeric()
    jyear <- 2021-as.numeric(jvaccmonth>jsurveymonth)
    reportedj1early <- paste0(jyear,"-",jvaccmonth,"-01") %>% as.Date(origin="1970-01-01")
    reportedj1late <- reportedj1early %>% add.months(1) %>% `-`(1)
    
    # check to see if the reported month agrees with the timing of daily surveys
    surveywindow1 <- j1$time_daily_survey[1] %>% as.character() %>% substr(1,8) %>% paste0("01") %>% as.Date(origin="1970-01-01")
    surveywindow2 <- j1$time_daily_survey[2] %>% as.character() %>% substr(1,8) %>% paste0("01") %>% as.Date(origin="1970-01-01")
    surveywindow <- seq.Date(from=surveywindow1,to=surveywindow2,by="1 month")
    if (reportedj1early %in% surveywindow) {
      kearly1 <- max(kearly,reportedj1early) %>% as.character()
      klate1 <-  min(klate,reportedj1late) %>% as.character()
      k <- c(i,kearly1,klate1)
    } else {
      print(paste0("i = ",i," reported month error"))
    }
  } else { # for people who reported a first shot immediately after the amendment was created or when they first enrolled
    jvaccmonth <- j$covid_vacc_first_m[1] %>% as.numeric()
    jsurveymonth <- j$time_daily_survey[1] %>% substr(6,7) %>% as.numeric()
    jyear <- 2021-as.numeric(jvaccmonth>jsurveymonth)
    
    kearly <- paste0(jyear,"-",jvaccmonth,"-01") %>% as.Date(origin="1970-01-01")
    klate <- kearly %>% add.months(1) %>% `-`(1)
    ksurvtime <- j$time_daily_survey[1] %>% substr(1,10) %>% as.Date(origin="1970-01-01") # going by daily survey datestamp
    
    klate1 <- min(klate,ksurvtime) %>% as.character()
    k <- c(i,as.character(kearly),klate1)
  }
  names(k) <- c("subject_id","vaccdate1_1","vaccdate1_2")
  vaccdtdates <- vaccdtdates %>% bind_rows(k)
}
vaccdtdates$vaccdate1_1 <- vaccdtdates$vaccdate1_1 %>% as.Date(origin="1970-01-01")
vaccdtdates$vaccdate1_2 <- vaccdtdates$vaccdate1_2 %>% as.Date(origin="1970-01-01")
vaccdtdates$window <- as.numeric(vaccdtdates$vaccdate1_2-vaccdtdates$vaccdate1_1)
vaccdtdates$window %>% as.factor() %>% summary()
vaccdtdates %>% filter(window<0) %>% nrow()

# figure out vaccine reported dates from rcmain
vacc1mainsubjs <- vaccmain %>% filter(covid_shot_yn=='1' | covid_shot_yn_recent=='1') %>% select(subject_id) %>% unlist() %>% unique() %>% as.numeric() %>% sort()
vacc1mainhist  <- vaccmain %>% filter(subject_id %in% vacc1mainsubjs)
vaccmaindates <-data.frame()
for (i in vacc1mainsubjs) {
  j0 <- vacc1mainhist %>% filter(subject_id==i)
  j1 <- j0 %>% select(vacc_time,covid_shot_yn,covid_shot_m)
  j2 <- j0 %>% select(vacc_time_recent,covid_shot_yn_recent,covid_shot_m_recent) %>% rename(vacc_time=vacc_time_recent,covid_shot_yn=covid_shot_yn_recent,covid_shot_m=covid_shot_m_recent)
  j <- bind_rows(j1,j2) %>% filter(!is.na(vacc_time))
  j$vacc_time <- j$vacc_time %>% substr(1,10) %>% as.Date(origin="1970-01-01")
  j <- j %>% arrange(vacc_time)
  if (j$covid_shot_yn[1]=='1') { # if there's only one row in j, or if the first j is already 1
   # go by the month of the first or only row
   jvaccmonth <- j$covid_shot_m[1] %>% as.numeric()
   jsurveymonth <- j$vacc_time[1] %>% substr(6,7) %>% as.numeric()
   jyear <- 2021-as.numeric(jvaccmonth>jsurveymonth)
   kearly <- paste0(jyear,"-",jvaccmonth,"-01") %>% as.Date(origin="1970-01-01")
   klate <- kearly %>% add.months(1) %>% `-`(1)
   ksurvtime <- j$vacc_time[1] %>% as.Date(origin="1970-01-01") # going by daily survey datestamp
   klate1 <- min(klate,ksurvtime) %>% as.character()
   k <- c(i,as.character(kearly),klate1)
  } else {
   # possible vacc window from reported 
   kearly <- j$vacc_time[1]
   klate <- j$vacc_time[2]
   k <- c(i,as.character(kearly),as.character(klate))
   # check the reported month fits in the window
   jvaccmonth <- j$covid_shot_m[2] %>% as.numeric()
   jsurveymonth <- j$vacc_time[2] %>% substr(6,7) %>% as.numeric()
   jyear <- 2021-as.numeric(jvaccmonth>jsurveymonth)
   reportedj1early <- paste0(jyear,"-",jvaccmonth,"-01") %>% as.Date(origin="1970-01-01")
   reportedj1late <- reportedj1early %>% add.months(1) %>% `-`(1)
   
   # check to see if the reported month agrees with the timing of daily surveys
   surveywindow1 <- j$vacc_time[1] %>% as.character() %>% substr(1,8) %>% paste0("01") %>% as.Date(origin="1970-01-01")
   surveywindow2 <- j$vacc_time[2] %>% as.character() %>% substr(1,8) %>% paste0("01") %>% as.Date(origin="1970-01-01")
   surveywindow <- seq.Date(from=surveywindow1,to=surveywindow2,by="1 month")
   if (reportedj1early %in% surveywindow) {
     kearly1 <- max(kearly,reportedj1early) %>% as.character()
     klate1 <-  min(klate,reportedj1late) %>% as.character()
     k <- c(i,kearly1,klate1)
   } else {
     print(paste0("i = ",i," reported month error"))
   }
  }
  names(k) <- c("subject_id","vaccdate1_1","vaccdate1_2")
  vaccmaindates <- vaccmaindates %>% bind_rows(k)
}
vaccmaindates$vaccdate1_1 <- vaccmaindates$vaccdate1_1 %>% as.Date(origin="1970-01-01")
vaccmaindates$vaccdate1_2 <- vaccmaindates$vaccdate1_2 %>% as.Date(origin="1970-01-01")
vaccdates_firstshot <- bind_rows(vaccdtdates,vaccmaindates) %>% arrange(subject_id %>% as.numeric())

# dedupe and combine dates from both sorts. Info from the daily survey will probably be more accurate
vaccdupesubjs <- vaccdates_firstshot$subject_id %>% subset(duplicated(.)) %>% unique() %>% as.numeric() %>% sort()
print(paste0("The following subjects have conflicting vacc1 dates: ",vaccdupesubjs %>% paste(collapse=" ")))
for (i in vaccdupesubjs) {
  j <- vaccdates_firstshot %>% filter(subject_id==i)
  j1 <- vaccdtdates %>% filter(subject_id==i)
  vaccdates_firstshot <- vaccdates_firstshot %>% filter(subject_id!=i)
  vaccdates_firstshot <- vaccdates_firstshot %>% bind_rows(j1)
}

vaccdates_firstshot <- vaccdates_firstshot %>% arrange(subject_id %>% as.numeric()) %>% select(subject_id,vaccdate1_1,vaccdate1_2)
summary(vaccdates_firstshot)

summary(vaccdates_firstshot$vaccdate1_1)
summary(vaccdates_firstshot$vaccdate1_2)

demohist <- demohist %>% left_join(vaccdates_firstshot,by=c("study_id"="subject_id"))

head(demohist[,c("study_id","vaccdate1_1","vaccdate1_2")])

```
serology data
Identify subjects that seroconverted: did the screen result change from first timepoint to second timepoint?

Categories:
0 = No conversion: ab- at enrollment and remained ab- at repeat testing
1 = Yes conversion: ab- at enrollment and then became ab+ at repeat testing
2 = Reversion: ab+ at enrollment and changed to ab- at repeat testing
3 = ab+ at enrollment and at repeat testing. We expect most ab+ at enrollment to remain ab+ at repeat testing.

Also, sero.diff = Number of days between first blood collection (at enrollment) and 2nd collection.
```{r serology}
sero1 <- read.csv("../source_data/serology_results_latest.csv",stringsAsFactors=F)
sero1 <- sero1 %>% filter(!is.na(subject_id)) %>%
                select(subject_id, collection_date, screen_result, igg_titer,igm_titer) %>%
                group_by(subject_id) %>%
                mutate(serum1=min(as.Date(as.character(collection_date)))) %>%
                mutate(serum2=max(as.Date(as.character(collection_date)))) %>%
                mutate(sero_daydiff=as.numeric(serum2-serum1))%>%
                mutate(seroconvert=NA)
sero1$collection_date <- sero1$collection_date %>% as.Date(origin="1970-01-01")
sero1 <- sero1 %>% arrange(as.numeric(subject_id),collection_date)
sero1 <- sero1 %>% ungroup()

sero1subjs <- sero1$subject_id %>% subset(duplicated(.)) %>% unique() %>% sort() # only people with more than one timepoint

sero1subjstatus <- sero1subjs %>% sapply(function(x) {sero1 %>% filter(subject_id==x) %>% arrange(collection_date) %>% pull(screen_result) %>% head(2) %>% paste(collapse=" ")}) # head(2) = only take the first two timepoints, if for whatever reason there are more than 2

names(sero1subjstatus) <- sero1subjs
sero1$seroconvert[sero1$subject_id %in% (sero1subjstatus %>% subset(.=="Negative Negative") %>% names())] <- 0
sero1$seroconvert[sero1$subject_id %in% (sero1subjstatus %>% subset(.=="Negative Positive") %>% names())] <- 1
sero1$seroconvert[sero1$subject_id %in% (sero1subjstatus %>% subset(.=="Positive Negative") %>% names())] <- 2
sero1$seroconvert[sero1$subject_id %in% (sero1subjstatus %>% subset(.=="Positive Positive") %>% names())] <- 3
sero1$seroconvert[is.na(sero1$seroconvert)] <- 9
sero1 <- sero1 %>% rename(ab_screen=screen_result,study_id=subject_id,date_visit=collection_date)

sero1$igg_titer[sero1$igg_titer==""] <- "Negative"
sero1$igm_titer[sero1$igm_titer==""] <- "Negative"
sero1$igg_titer[sero1$igg_titer=="Neg"] <- "Negative"

summary(sero1)
# seroconvert can go to demohist
seroconvert <- sero1 %>% select(study_id,seroconvert) %>% unique()
seroconvert$study_id <- seroconvert$study_id %>% as.character()
demohist <- demohist %>% left_join(seroconvert, by="study_id")
```
seropositivity at enrollment
```{r seropositive at enrollment}
allsubjs <- demohist$study_id %>% unique() %>% as.numeric() %>% sort() %>% as.character()
allsubjsenrollmentdate <- allsubjs %>% lapply(function(x) {pcr_screen %>% filter(study_id==x & !grepl("driveby",redcap_event_name)) %>% pull(sample_date)}) # enrollment date here refers to first blood sampling, so weekly saliva samples are excluded

allsubjsenrollmentdate <- allsubjsenrollmentdate %>% lapply(function(x) {if (length(x)>0) {min(x) %>% as.Date(origin="1970-01-01")} else {NA}}) # if a person never gave blood, they would get NA here instead of a date
names(allsubjsenrollmentdate) <- allsubjs

allsubjs_screenresult <- allsubjs  %>% lapply(function(x) {
  if (!is.na(allsubjsenrollmentdate[[x]])) {
    sero1 %>% filter(study_id==x & date_visit==allsubjsenrollmentdate[[x]]) %>% pull(ab_screen)
  } else {
    "never gave blood"
  }
})

# if a person didn't give blood, they would get "never gave blood" here. If the serology at the enrollment date wasn't done, then selecting (ab_screen) would give character(0). 
# In the next line that is converted to "unknown"

allsubjs_screenresult <- allsubjs_screenresult %>% lapply(function(x) {if (length(x)==0) {"unknown"} else {x}})
names(allsubjs_screenresult) <- allsubjs

# need to check if serology was done on a close date, if not necessarily the first date
unkserosubjs <- allsubjs %>% sapply(function(x) {allsubjs_screenresult[[x]]=="unknown"}) %>% subset(.==T) %>% names()
unkserosubjs
unkserosubjs %>% subset(. %in% sero1$study_id)

# 484 didn't give blood on the first day, but did give blood on the second day, and no other - use this
allsubjs_screenresult[["484"]] <- sero1 %>% filter(study_id==484) %>% pull(ab_screen)

# enrollment date should be modified for those people who gave blood closer to the date of their first g-ii sample
# determine whose blood got assayed more than once before follow-up

# attach redcap event name to sample date

sero2 <- sero1
sero2$helper <- paste(sero2$study_id,sero2$date_visit)
samplesevents <- allsamples %>% mutate(helper=paste(study_id,sample_date)) %>% select(helper,redcap_event_name) %>% filter(!grepl("driveby",redcap_event_name)) %>% unique()
sero2 <- sero2 %>% left_join(samplesevents,by="helper")

sero2a <- sero2 %>% filter(!grepl("ir",redcap_event_name))
sero2asubjs <- sero2a$study_id %>% unique()
sero2asubjcount <- sero2asubjs %>% sapply(function(x) {sero2a %>% filter(study_id==x) %>% nrow()})
names(sero2asubjcount) <- sero2asubjs
multibloodsubjs <- sero2asubjcount %>% subset(.>1) %>% names()
multibloodscreenresults <- multibloodsubjs %>% sapply(function(x) {allsubjs_screenresult[[x]]})
sero2a <- sero2 %>% filter((study_id %in% multibloodsubjs) & !grepl("ir",redcap_event_name))

# subjs whose serology status upon repeated testing
serochangesubjs <- multibloodsubjs %>% sapply(function(x) {sero2a %>% filter(study_id==x) %>% pull(ab_screen) %>% unique()}) %>% subset(lapply(.,length)>1) %>% names()

# 255 is not standard because of multi shedding visits
# 286 is a change

allsubjs_screenresult[["286"]] <- "Positive"

demohist$pos_enrollment <- demohist$study_id %>% sapply(function(x) {allsubjs_screenresult[[x]]})
```
merge in culture data from Stuart
culture data is in "status"->"culture_status", can be positive or negative
```{r merge culture data}
culturelatest <- read.csv(file="../source_data/culture_results_merged_latest.csv",stringsAsFactors = F)
culturelatest <- culturelatest %>% select(sample_id,subject_id,date_collect,sample_type,status) %>% rename(culture_status=status)

# fix culturelatest sample types
culturelatest$sample_type <- culturelatest$sample_type %>% gsub(" Derivative","",.)
culturelatest$sample_type[culturelatest$sample_type=="G2 Fine Aerosol Wash"] <- "G-II Fine Aerosol"
culturelatest$sample_type[culturelatest$sample_type=="G2 5-Micron Impactor"] <- "G-II Coarse Aerosol"
culturelatest$sample_type[culturelatest$sample_type=="IcePac Sample"] <- "IcePac"

# figure which culture samples went with masks
mc1 <- masks_cough
mc1$allcodes <- mc1[,c("fine_aero_barcode","fine_children","coarse_barcode","coarse_children","cone_barcode","cone_children")] %>% apply(1,function(x) {x %>% na.omit() %>% paste(collapse=";")})
culturelatestmaskyn <- culturelatest$sample_id %>% sapply(function(x) {mc1 %>% filter(grepl(x,allcodes)) %>% pull(mask_yn)})
culturelatest$maskyn <- culturelatestmaskyn %>% lapply(function(x) {if (length(x)!=0) {x} else {""}}) %>% unlist() # character(0) to text
culturelatest$sample_type[culturelatest$maskyn=="1"] <- culturelatest$sample_type[culturelatest$maskyn=="1"] %>% paste0(" mask")

# bind culturelatest by participant, day, and sample type
culturelatest <- culturelatest %>% mutate(helper=paste(subject_id,date_collect,sample_type))

qpcr_allsamples$sample_type[qpcr_allsamples$sample_type=="G-II Fine Aerosol Wash"] <- "G-II Fine Aerosol"
qpcr_allsamples$sample_type[qpcr_allsamples$sample_type=="G-II Fine Aerosol Wash mask"] <- "G-II Fine Aerosol mask"
qpcr_allsamples$sample_type[qpcr_allsamples$sample_type=="G-II 5-Micron Impactor"] <- "G-II Coarse Aerosol"
qpcr_allsamples$sample_type[qpcr_allsamples$sample_type=="G-II 5-Micron Impactor mask"] <- "G-II Coarse Aerosol mask"
qpcr_allsamples$sample_type[qpcr_allsamples$sample_type=="IcePac Sample"] <- "IcePac"

culture_status_only <- culturelatest %>% select(culture_status,helper) %>% unique()
culture_status_only$helper %>% subset(duplicated(.))
# should be character(0)

qpcr_allsamples <- qpcr_allsamples %>% mutate(helper=paste(study_id,sample_date,sample_type))
qpcr_allsamples <- qpcr_allsamples %>% left_join(culture_status_only,by="helper")
qpcr_allsamples$helper <- NULL
qpcr_allsamples$culture_status[is.na(qpcr_allsamples$culture_status)] <- "not cultured"
summary(qpcr_allsamples$culture_status %>% as.factor())
```
assign classes
```{r class assignments}
# what classes are supposed to be
demohistclasses <- list()
demohistclasses[["character"]] <- c("study_id","other_race","race","onset_from_sx_card","pos_enrollment","vacc_status")
demohistclasses[["numeric"]] <- c("age_entry","alc_week","drinks_p_day","drinks_p_week","bmi",all_sx_ret,"ur_ret","lr_ret","sys_ret","gi_ret")
demohistclasses[["factor"]] <- c("sex_birth","cigarette_ever","singing_lessons","singing_ensemble","latino","tag","taste_smell_yn_covid_ret","sx_med_care_ret","seroconvert")
demohistclasses[["logical"]] <- c("cri","multiracial","fever_ever")
demohistclasses[["Date"]] <- c("real_onset","vaccdate1_1","vaccdate1_2","date_of_enrollment")

pcrscreenclasses <- list()
pcrscreenclasses[["character"]] <- c("tac_id","study_id","redcap_event_name","redcap_repeat_instance","sample_id","sample_type")
pcrscreenclasses[["Date"]] <- c("sample_date")
pcrscreenclasses[["time"]]<- c("symptom_time")
pcrscreenclasses[["numeric"]] <- c("ct_sgene_screen","ct_orf1ab_screen","ct_ngene_screen","oral_temp",ur_sx,lr_sx,gi_sx,sys_sx,"ur","lr","gi","sys","ox_sat","days_past_onset","ct_avg_screen")
pcrscreenclasses[["factor"]] <- c("taste_smell_loss")
pcrscreenclasses[["logical"]] <- c("fever_samp")

qpcrclasses <- list()
qpcrclasses[["character"]]<- c("study_id","exp_name","sample_id","sample_type")
qpcrclasses[["numeric"]]<- c("ct_ngene_qpcr","ct_sgene_qpcr","ct_orfgene_qpcr","rxn_quant","sample_quant","cough_count","reserv_time","gii_reps","cens","av_quant","mdfy_sample_quant")
qpcrclasses[["Date"]]<- c("sample_date","exp_date")
qpcrclasses[["factor"]]<- c("gii_sessnum","mask_yn","mask_type","nd_lod_loq","gii","culture_status")
qpcrclasses[["logical"]]<- c("S.gene.dropout","any_rep_lod1","all_rep_lod2","all_rep_loq","loq")

sero1classes <- list()
sero1classes[["character"]]<- c("study_id")
sero1classes[["numeric"]]<- c("sero_daydiff")
sero1classes[["Date"]]<- c("date_visit","serum1","serum2")
sero1classes[["factor"]]<- c("igg_titer","igm_titer","seroconvert")
sero1classes[["logical"]]<- c("ab_screen")

# fix things here
for (i in demohistclasses[["factor"]]) {
  demohist[,i] <- demohist[,i] %>% as.factor()
}
demohistnumsublist <- c("alc_week","drinks_p_day","bmi")
for (i in demohistnumsublist) {
  demohist[,i] <- demohist[,i] %>% as.numeric()
}
demohist$cri <- demohist$cri %>% as.logical()
demohist$multiracial[demohist$multiracial=="Yes"] <- "TRUE"
demohist$multiracial[demohist$multiracial=="No"] <- "FALSE"
demohist$multiracial <- demohist$multiracial %>% as.logical()

pcr_screen$symptom_time <- pcr_screen$symptom_time %>% as.POSIXct(tz="America/New_York")
pcr_screen$taste_smell_loss <- pcr_screen$taste_smell_loss %>% as.factor()
pcr_screen$ox_sat <- pcr_screen$ox_sat %>% as.numeric()
pcr_screen$fever_samp <- pcr_screen$fever_samp %>% as.logical()

qpcr_allsamples$sample_date <- qpcr_allsamples$sample_date %>% as.Date(origin="1970-01-01")
qpcr_allsamples$gii_sessnum <- qpcr_allsamples$gii_sessnum %>% as.factor()
qpcr_allsamples$mask_yn <- qpcr_allsamples$mask_yn %>% as.factor()
qpcr_allsamples$mask_type <- qpcr_allsamples$mask_type %>% as.factor()
qpcr_allsamples$cough_count <- qpcr_allsamples$cough_count %>% as.numeric()
qpcr_allsamples$reserv_time <- qpcr_allsamples$reserv_time %>% as.numeric()
qpcr_allsamples$nd_lod_loq <- qpcr_allsamples$nd_lod_loq %>% as.factor()
qpcr_allsamples$gii <- qpcr_allsamples$gii %>% as.factor()
qpcr_allsamples$culture_status <- qpcr_allsamples$culture_status %>% as.factor()

sero1 <- sero1 %>% as.data.frame()
for (i in sero1classes[["factor"]]) {
  sero1[,i] <- sero1[,i] %>% as.factor()
}
sero1$ab_screen[sero1$ab_screen=="Negative"] <- "FALSE"
sero1$ab_screen[sero1$ab_screen=="Positive"] <- "TRUE"
sero1$ab_screen <- sero1$ab_screen %>% as.logical()
sero1$study_id <- sero1$study_id %>% as.character()
```
check classes of columns
```{r check classes}
demohistclassesassigned <- names(demohistclasses) %>% lapply(function(x) {rep(x,length(demohistclasses[[x]]))})
for (i in 1:length(demohistclasses)) {
  names(demohistclassesassigned[[i]]) <- demohistclasses[[i]]
}
demohistclassesassigned <- demohistclassesassigned %>% unlist()
demohistclassesactual <- colnames(demohist) %>% sapply(function(x) {demohist[,x] %>% class()})

pcrscreenclassesassigned <- names(pcrscreenclasses) %>% lapply(function(x) {rep(x,length(pcrscreenclasses[[x]]))})
for (i in 1:length(pcrscreenclasses)) {
  names(pcrscreenclassesassigned[[i]]) <- pcrscreenclasses[[i]]
}
pcrscreenclassesassigned <- pcrscreenclassesassigned %>% unlist()
pcrscreenclassesactual <- colnames(pcr_screen) %>% sapply(function(x) {pcr_screen[,x] %>% class()})

qpcrclassesassigned <- names(qpcrclasses) %>% lapply(function(x) {rep(x,length(qpcrclasses[[x]]))})
for (i in 1:length(qpcrclasses)) {
  names(qpcrclassesassigned[[i]]) <- qpcrclasses[[i]]
}
qpcrclassesassigned <- qpcrclassesassigned %>% unlist()
qpcr_allsamples1 <- as.data.frame(qpcr_allsamples) # get around some tibble properties
qpcrclassesactual <- colnames(qpcr_allsamples1) %>% sapply(function(x) {qpcr_allsamples1[,x] %>% class()})

sero1classesassigned <- names(sero1classes) %>% lapply(function(x) {rep(x,length(sero1classes[[x]]))})
for (i in 1:length(sero1classes)) {
  names(sero1classesassigned[[i]]) <- sero1classes[[i]]
}
sero1classesassigned <- sero1classesassigned %>% unlist()
sero1classesactual <- colnames(sero1) %>% sapply(function(x) {sero1[,x] %>% unlist() %>% class()})

allclasses <- c("demohist","pcrscreen","qpcr","sero1")

for (q in allclasses) {
  assignedclasses <- eval(parse(text=paste0(q,"classesassigned")))
  actuallist <- eval(parse(text=paste0(q,"classesactual")))
  for (p in names(actuallist)) {
    pactual <- actuallist[p] %>% gsub("integer","numeric",.)
    if (pactual!=assignedclasses[p]) {
      print(paste(q,":",p,actuallist[p],assignedclasses[p]))
    }
  }
}
#stop()
```
Write to directory the pcr plus covariables (merged) data
Add a datestamp
```{r write list of dataframes}
cov_pcr_sera <- list(demohist,pcr_screen,qpcr_allsamples,sero1)
names(cov_pcr_sera) <- c("demohist","pcr_screen","qpcr_allsamples","sero1")
rdsfilename <- paste0("cov_pcr_sera_",Sys.Date() %>% format("%Y%m%d"),".RDS")
saveRDS(cov_pcr_sera, file = rdsfilename)
```
