---
title: "Fixed fine and coarse"
author: "Jianyu"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
#library(stringr)
#library(readr)
library(tidyverse)
#library(lubridate)
library(forcats)
library(lmec)
library(kableExtra)
library(lme4)
source("z_create_matrix.R")
```

#### All samples
#import data
```{r}
dF<-readRDS("../working_files/nonpaired/nonpaired_pcr_dat_negsero.RDS") 
dF <- dF %>%  mutate(mdfy_sample_quant = if_else(cens==1, 75, mdfy_sample_quant)) 
pcr_dat_clean<-dF %>% mutate(ln_final_copies=log(mdfy_sample_quant))
pcrdatclean_sampletypes <- pcr_dat_clean$sample_type %>% unique()

sarslist <- pcrdatclean_sampletypes %>% lapply(function(x) {pcr_dat_clean %>% filter(sample_type==x)})
names(sarslist) <- pcrdatclean_sampletypes
```

###GM and GSD for G-II Fine Aerosol

####using lmer
```{r}
lmer_fine <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist[["G-II Fine Aerosol"]])
summary(lmer_fine)


beta_fine <- fixef(lmer_fine)
sigma_fine <- sqrt(diag(vcov(lmer_fine)))
low_ci_fold_inc <- signif(exp(beta_fine - 1.96*(sigma_fine)), 2)
up_ci_fold_inc <- signif(exp(beta_fine + 1.96*(sigma_fine)), 2)
fold_inc <- signif(exp(beta_fine), 2)
fold_inc_var_fine <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var_fine  
```

####using lmec
```{r}
sarsdataset <- sarslist[["G-II Fine Aerosol"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```

###GM and GSD for G-II Fine Aerosol with mask
####using lmec
###MLmcmc
```{r}
sarsdataset <- sarslist[["G-II Fine Aerosol mask"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```

###GM and GSD for G-II Coarse Aerosol

####using lmer
```{r}
lmer <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist[["G-II Coarse Aerosol"]])
summary(lmer)


beta <- fixef(lmer)
sigma <- sqrt(diag(vcov(lmer)))
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var  
```

####using lmec
```{r}
sarsdataset <- sarslist[["G-II Coarse Aerosol"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```


#### Other variants
#import data
```{r}
dF_other <- pcr_dat_clean %>% filter(s_gen_pos_id == "TRUE")

pcrdatclean_sampletypes <- dF_other$sample_type %>% unique()

sarslist_other <- pcrdatclean_sampletypes %>% lapply(function(x) {dF_other %>% filter(sample_type==x)})
names(sarslist_other) <- pcrdatclean_sampletypes
```

###GM and GSD for G-II Fine Aerosol

####using lmer
```{r}
lmer_fine <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist_other[["G-II Fine Aerosol"]])
summary(lmer_fine)


beta_fine <- fixef(lmer_fine)
sigma_fine <- sqrt(diag(vcov(lmer_fine)))
low_ci_fold_inc <- signif(exp(beta_fine - 1.96*(sigma_fine)), 2)
up_ci_fold_inc <- signif(exp(beta_fine + 1.96*(sigma_fine)), 2)
fold_inc <- signif(exp(beta_fine), 2)
fold_inc_var_fine <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var_fine  
```
####using lmec
```{r}
sarsdataset <- sarslist_other[["G-II Fine Aerosol"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```

###GM and GSD for G-II Fine Aerosol mask
####using lmec
```{r}
sarsdataset <- sarslist_other[["G-II Fine Aerosol mask"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```
###GM and GSD for G-II Coarse Aerosol

####using lmer
```{r}
lmer <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist_other[["G-II Coarse Aerosol"]])
summary(lmer)


beta <- fixef(lmer)
sigma <- sqrt(diag(vcov(lmer)))
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var   
```
####using lmec
```{r}
sarsdataset <- sarslist_other[["G-II Coarse Aerosol"]]
  Z_current <- z_create_matrix(sarsdataset)
  X <- matrix(rep(1, length(sarsdataset$ln_final_copies)), ncol=1) #for intercept only
  cluster <- as.numeric(as.factor(sarsdataset$study_id))
 fit_ML <- lmec(yL = sarsdataset$ln_final_copies,
                  cens = sarsdataset$cens,
                  X = X, 
                  Z = Z_current,
                  cluster=cluster,
                  method='ML',
                  maxstep=200)
beta <- fit_ML$beta[1]
sigma <- sqrt(fit_ML$varFix[1])
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind(var, fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var
```

###alpha
```{r}
dF_alpha <- pcr_dat_clean %>% filter(s_gen_pos_id == "FALSE")

pcrdatclean_sampletypes <- dF_alpha$sample_type %>% unique()

sarslist_alpha <- pcrdatclean_sampletypes %>% lapply(function(x) {dF_alpha %>% filter(sample_type==x)})
names(sarslist_alpha) <- pcrdatclean_sampletypes

```
###GM and GSD for G-II Fine Aerosol
####using lmer
```{r}
lmer_fine <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist_alpha[["G-II Fine Aerosol"]])
summary(lmer_fine)


beta_fine <- fixef(lmer_fine)
sigma_fine <- sqrt(diag(vcov(lmer_fine)))
low_ci_fold_inc <- signif(exp(beta_fine - 1.96*(sigma_fine)), 2)
up_ci_fold_inc <- signif(exp(beta_fine + 1.96*(sigma_fine)), 2)
fold_inc <- signif(exp(beta_fine), 2)
fold_inc_var_fine <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var_fine  
```
###GM and GSD for G-II Coarse Aerosol
####using lmer
```{r}
lmer <- lmer (ln_final_copies ~ (1|study_id) + (1|study_id:sample_date) , data=sarslist_alpha[["G-II Coarse Aerosol"]])
summary(lmer)

beta <- fixef(lmer)
sigma <- sqrt(diag(vcov(lmer)))
low_ci_fold_inc <- signif(exp(beta - 1.96*(sigma)), 2)
up_ci_fold_inc <- signif(exp(beta + 1.96*(sigma)), 2)
fold_inc <- signif(exp(beta), 2)
fold_inc_var <- cbind("intercept", fold_inc, low_ci_fold_inc, up_ci_fold_inc)
fold_inc_var   
```